### 2023-11-22 paper.tex
Put together template

\documentclass{article}

\usepackage[final]{assets/techreport}
%\usepackage[nonatbib]{assets/techreport}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\title{VitalLens 1.0 Technical Report}

\author{%
  Philipp V.~Rouast \\
  Rouast Labs\\
  \texttt{philipp@rouast.com} \\
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

\section{Introduction}

Please read the instructions below carefully and follow them faithfully. \textbf{Important:} This year the checklist will be submitted separately from the main paper in OpenReview, please review it well ahead of the submission deadline: \url{https://neurips.cc/public/guides/PaperChecklist}.

\subsection{Style}

Papers to be submitted to NeurIPS 2023 must be prepared according to the
instructions presented here. Papers may only be up to {\bf nine} pages long,
including figures. Additional pages \emph{containing only acknowledgments and
references} are allowed. Papers that exceed the page limit will not be
reviewed, or in any other way considered for presentation at the conference.

The margins in 2023 are the same as those in previous years.

Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
NeurIPS website as indicated below. Please make sure you use the current files
and not previous versions. Tweaking the style files may be grounds for
rejection.

\subsection{Retrieval of style files}

The style files for NeurIPS and other conference information are available on
the website at
\begin{center}
  \url{http://www.neurips.cc/}
\end{center}
The file \verb+neurips_2023.pdf+ contains these instructions and illustrates the
various formatting requirements your NeurIPS paper must satisfy.

The only supported style file for NeurIPS 2023 is \verb+neurips_2023.sty+,
rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
  Microsoft Word, and RTF are no longer supported!}

The \LaTeX{} style file contains three optional arguments: \verb+final+, which
creates a camera-ready copy, \verb+preprint+, which creates a preprint for
submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
\verb+natbib+ package for you in case of package clash.

\paragraph{Preprint option}
If you wish to post a preprint of your work online, e.g., on arXiv, using the
NeurIPS style, please use the \verb+preprint+ option. This will create a
nonanonymized version of your work with the text ``Preprint. Work in progress.''
in the footer. This version may be distributed as you see fit, as long as you do not say which conference it was submitted to. Please \textbf{do
  not} use the \verb+final+ option, which should \textbf{only} be used for
papers accepted to NeurIPS.

At submission time, please omit the \verb+final+ and \verb+preprint+
options. This will anonymize your submission and add line numbers to aid
review. Please do \emph{not} refer to these line numbers in your paper as they
will be removed during generation of camera-ready copies.

The file \verb+neurips_2023.tex+ may be used as a ``shell'' for writing your
paper. All you have to do is replace the author, title, abstract, and text of
the paper with your own.

The formatting instructions contained in these style files are summarized in
Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

\section{Scope and Limitations of this Technical Report}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

For the final version, authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Datasets}
\label{headings}

All headings should be lower case (except for first word and proper nouns),
flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the heading in
bold, flush left, and inline with the text, with the heading followed by 1\,em
of space.

\section{Methodology}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.  Citations may be
author/year or numeric, as long as you maintain internal consistency.  As to the
format of the references themselves, any style is acceptable as long as it is
used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations appropriate for
use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may add the
following before loading the \verb+neurips_2023+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add the optional
argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{neurips_2023}
\end{verbatim}

As submission is double blind, refer to your own published work in the third
person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
previous work [4].'' If you cite your other papers that are not widely available
(e.g., a journal paper under review), use anonymous author names in the
citation, e.g., an author of the form ``A.\ Anonymous'' and include a copy of the anonymized paper in the supplementary material.

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote, indicate
footnotes with a number\footnote{Sample of the first footnote.} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table number and
title always appear before the table.  See Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical rules.} We
strongly suggest the use of the \verb+booktabs+ package, which allows for
typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Math}
Note that display math in bare TeX commands will not create correct line numbers for submission. Please use LaTeX (or AMSTeX) commands for unnumbered display math. (You really shouldn't be using \$\$ anyway; see \url{https://tex.stackexchange.com/questions/503/why-is-preferable-to} and \url{https://tex.stackexchange.com/questions/40492/what-are-the-differences-between-align-equation-and-displaymath} for more information.)

\subsection{Final instructions}

Do not change any aspects of the formatting parameters in the style files.  In
particular, do not modify the width or length of the rectangle the text should
fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Results and Discussion}

Please prepare submission files with paper size ``US Letter,'' and not, for
example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file must only
contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
  available out-of-the-box on most Linux machines.

\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
  "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
  the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
figure width as a multiple of the line width as in the example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
necessary.

\begin{ack}
Use unnumbered first level headings for the acknowledgments. All acknowledgments
go at the end of the paper before the list of references. Moreover, you are required to declare
funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2023/PaperInformation/FundingDisclosure}.

Do {\bf not} include this section in the anonymized submission, only in the final paper. You can use the \texttt{ack} environment provided in the style file to autmoatically hide this section in the anonymized submission.
\end{ack}

\section{Conclusion}

Authors may wish to optionally include extra information (complete proofs, additional experiments and plots) in the appendix. All such materials should be part of the supplemental material (submitted separately) and should NOT be included in the main submission.

\section*{References}

References follow the acknowledgments in the camera-ready paper. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references.
Note that the Reference section does not count towards the page limit.
\medskip

{
\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2023-12-05 paper.tex
Start writing

@@
 \begin{abstract}
-  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
-  both the left- and right-hand margins. Use 10~point type, with a vertical
-  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
-  bold, and in point size 12. Two line spaces precede the abstract. The abstract
-  must be limited to one paragraph.
+We report the development of VitalLens, an app that estimates vital signs such as heart rate and respiration rate from selfie video in real time.
+The estimation engine of VitalLens is a computer vision model trained on a diverse dataset of video and gold-standard ground truth physiological sensor data.
+% TODO: Summarize findings
 \end{abstract}
@@
 
-Please read the instructions below carefully and follow them faithfully.
-\textbf{Important:} This year the checklist will be submitted separately from the main paper in OpenReview, please review it well ahead of the submission deadline: \url{https://neurips.cc/public/guides/PaperChecklist}.
+Video of the human face and upper body contains signals with information about the subject's physiological state.
+Given appropriate circumstances, these signals can be strong enough to estimate vital signs such as heart rate and respiratory rate.
+This process is known as Remote Photoplethysmography (rPPG); many rPPG approaches have been proposed, including handcrafted algorithms and ones learned from empirical data.
+Depending on the intended application, we can measure advantages and drawbacks of rPPG approaches in the accuracy, inference speed, and robustness regarding various factors impacting estimation performance.
 
-\subsection{Style}
+Paragraph on G, CHROM, and POS.
 
-Papers to be submitted to NeurIPS 2023 must be prepared according to the
-instructions presented here. Papers may only be up to {\bf nine} pages long,
-including figures. Additional pages \emph{containing only acknowledgments and
-references} are allowed. Papers that exceed the page limit will not be
-reviewed, or in any other way considered for presentation at the conference.
+Paragraph on DeepPhys and MTTS-CAN.
 
-The margins in 2023 are the same as those in previous years.
-
-Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
-NeurIPS website as indicated below. Please make sure you use the current files
-and not previous versions. Tweaking the style files may be grounds for
-rejection.
-
-\subsection{Retrieval of style files}
-
-The style files for NeurIPS and other conference information are available on
-the website at
-\begin{center}
-  \url{http://www.neurips.cc/}
-\end{center}
-The file \verb+neurips_2023.pdf+ contains these instructions and illustrates the
-various formatting requirements your NeurIPS paper must satisfy.
-
-The only supported style file for NeurIPS 2023 is \verb+neurips_2023.sty+,
-rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
-  Microsoft Word, and RTF are no longer supported!}
-
-The \LaTeX{} style file contains three optional arguments: \verb+final+, which
-creates a camera-ready copy, \verb+preprint+, which creates a preprint for
-submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
-\verb+natbib+ package for you in case of package clash.
-
-\paragraph{Preprint option}
-If you wish to post a preprint of your work online, e.g., on arXiv, using the
-NeurIPS style, please use the \verb+preprint+ option. This will create a
-nonanonymized version of your work with the text ``Preprint. Work in progress.''
-in the footer. This version may be distributed as you see fit, as long as you do not say which conference it was submitted to. Please \textbf{do
-  not} use the \verb+final+ option, which should \textbf{only} be used for
-papers accepted to NeurIPS.
-
-At submission time, please omit the \verb+final+ and \verb+preprint+
-options. This will anonymize your submission and add line numbers to aid
-review. Please do \emph{not} refer to these line numbers in your paper as they
-will be removed during generation of camera-ready copies.
-
-The file \verb+neurips_2023.tex+ may be used as a ``shell'' for writing your
-paper. All you have to do is replace the author, title, abstract, and text of
-the paper with your own.
-
-The formatting instructions contained in these style files are summarized in
-Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.
+Acknowledge other models we don't compare because lack focus on real-time.
 
@@
 
-The text must be confined within a rectangle 5.5~inches (33~picas) wide and
-9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
-type with a vertical spacing (leading) of 11~points.  Times New Roman is the
-preferred typeface throughout, and will be selected for you by default.
-Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
-indentation.
+This report focuses on the capabilities and limitations of VitalLens.
+The estimation engine of VitalLens is a recurrent convolutional model trained on a diverse dataset of video and gold-standard ground truth physiological data.
 
-The paper title should be 17~point, initial caps/lower case, bold, centered
-between two horizontal rules. The top rule should be 4~points thick and the
-bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
-below the title to rules. All pages should start at 1~inch (6~picas) from the
-top of the page.
-
-For the final version, authors' names are set in boldface, and each name is
-centered above the corresponding address. The lead author's name is to be listed
-first (left-most), and the co-authors' names (if different address) are set to
-follow. If there is only one co-author, list both author and co-author side by
-side.
-
-Please pay special attention to the instructions in Section \ref{others}
-regarding figures, tables, acknowledgments, and references.
+This report does not contain any further detail about the architecture of the model or training methodology -
+we reserve this for future publication.
 
@@
 
-All headings should be lower case (except for first word and proper nouns),
-flush left, and bold.
+Two major datasets are involved with training and evaluation of VitalLens:
+The \textit{PROSIT} dataset and the \textit{Vital Videos} dataset.
+The vast majority of the data used to train VitalLens comes from PROSIT.
+We enrich the training data with a subset of data from the Vital Videos dataset, as detailed in this section.
+The main Vital Videos dataset serves as a test set for the evaluation of VitalLens.
 
-First-level headings should be in 12-point type.
+\subsection{PROSIT}
 
-\subsection{Headings: second level}
+PROSIT (\underline{P}hysiological \underline{R}ecordings \underline{O}f \underline{S}ubjects using \underline{I}maging \underline{T}echnology) is our in-house dataset for practical rPPG applications.
 
-Second-level headings should be in 10-point type.
+\paragraph{Participant recruitment and session protocol.} 
+As part of our goal of creating a diverse rPPG dataset for practical applications, we recruit participants and collect data at various locations such as residential homes, offices, libraries, and clubs.
+Each potential participant was required to go through an informed consent process in accordance with Australian privacy laws prior to participation.
+During the session, participants are asked to complete tasks on an iPad while sensor data is being recorded.
 
-\subsubsection{Headings: third level}
+\paragraph{Sensors and collected data.}
+The time-synchronized sensor array used for PROSIT consists of a video camera, electrocardiogram (ECG), pulse oximetry, blood pressure monitor, and an ambient light sensor.
+This yields a rich set of data including video, ECG, PPG, SpO2, respiration, blood pressure, and ambient luminance.
+We also collect age, gender, height, and skin type metadata according to the Fitzpatrick scale.
 
-Third-level headings should be in 10-point type.
+\paragraph{Pre-processing.}
+We pre-process and split each session into small chunks of 5-20 seconds with valid video and signals.
+As part of this step, we also calculate summary vitals for each chunk from the continuous signals, and extract further metadata such as the amount of participant movement and illumination variation.
 
-\paragraph{Paragraphs}
+\paragraph{Dataset size and split.}
+Development of PROSIT is ongoing.
+As of the writing of this report, it comprises 157 unique participants across 173 recording sessions in 45 different locations.
+This results in a total of 9,767 chunks or 27.8 hours of data.
 
-There is also a \verb+\paragraph+ command available, which sets the heading in
-bold, flush left, and inline with the text, with the heading followed by 1\,em
-of space.
+\begin{table}[h!]
+  \caption{PROSIT Dataset Size}
+  \label{prosit-table}
+  \centering
+  \begin{tabular}{llll}
+    \toprule
+    Split      & \# Participants & \# Chunks & Time   \\
+    \midrule
+    Training   & 114 			 & 6,765     & 19.4 h \\
+    Validation & 23 			 & 1,599     & 4.5 h  \\
+    Test       & 20     			 & 1,403     & 3.9 h  \\
+    \midrule
+    Total      & 157			 & 9,767  	 & 27.8 h \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+Each participant is randomly assigned to be part of either the \textit{training}, \textit{validation}, or \textit{test} set to ensure that all participants seen during validation and test are previously unseen by the model.
+
+\paragraph{Participant diversity.}
+The skin type diversity of PROSIT is lacking, with too many type 2 and too few type 5 and 6 participants.
+This is why we additionally use the Vital Videos Ghana training set when training our model.
+
+TODO: Histogram of ages x n participants
+TODO: Histogram of skin types x n participants
+TODO: Histogram of genders x n participants
+
+\paragraph{Vitals diversity.}
+There are also several participants who have an irregular heartbeat.
+
+TODO: Histogram of HR x n chunks
+TODO: Histogram of RR x n chunks
+TODO: Histogram of BP x n chunks
+TODO: Histogram of SpO2 x n chunks
+
+\subsection{Vital Videos}
+
+\subsubsection{Vital Videos Medium}
+
+Used for test.
+
+Pieter's dataset
+Number of unique subjects
+Total duration
+Statistics
+Distributions of metadata
+
+\subsubsection{Vital Videos Ghana}
+
+Used for training.
+
+Pieter's dataset
+Number of unique subjects
+Total duration
+Statistics
+Distributions of metadata
 
@@
 
-These instructions apply to everyone.
-
-\subsection{Citations within the text}
-
-The \verb+natbib+ package will be loaded for you by default.  Citations may be
-author/year or numeric, as long as you maintain internal consistency.  As to the
-format of the references themselves, any style is acceptable as long as it is
-used consistently.
-
-The documentation for \verb+natbib+ may be found at
-\begin{center}
-  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
-\end{center}
-Of note is the command \verb+\citet+, which produces citations appropriate for
-use in inline text.  For example,
-\begin{verbatim}
-   \citet{hasselmo} investigated\dots
-\end{verbatim}
-produces
-\begin{quote}
-  Hasselmo, et al.\ (1995) investigated\dots
-\end{quote}
-
-If you wish to load the \verb+natbib+ package with options, you may add the
-following before loading the \verb+neurips_2023+ package:
-\begin{verbatim}
-   \PassOptionsToPackage{options}{natbib}
-\end{verbatim}
-
-If \verb+natbib+ clashes with another package you load, you can add the optional
-argument \verb+nonatbib+ when loading the style file:
-\begin{verbatim}
-   \usepackage[nonatbib]{neurips_2023}
-\end{verbatim}
-
-As submission is double blind, refer to your own published work in the third
-person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
-previous work [4].'' If you cite your other papers that are not widely available
-(e.g., a journal paper under review), use anonymous author names in the
-citation, e.g., an author of the form ``A.\ Anonymous'' and include a copy of the anonymized paper in the supplementary material.
-
-\subsection{Footnotes}
-
-Footnotes should be used sparingly.  If you do require a footnote, indicate
-footnotes with a number\footnote{Sample of the first footnote.} in the
-text. Place the footnotes at the bottom of the page on which they appear.
-Precede the footnote with a horizontal rule of 2~inches (12~picas).
-
-Note that footnotes are properly typeset \emph{after} punctuation
-marks.\footnote{As in this example.}
-
-\subsection{Figures}
-
-\begin{figure}
-  \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  \caption{Sample figure caption.}
-\end{figure}
-
-All artwork must be neat, clean, and legible. Lines should be dark enough for
-purposes of reproduction. The figure number and caption always appear after the
-figure. Place one line space before the figure caption and one line space after
-the figure. The figure caption should be lower case (except for first word and
-proper nouns); figures are numbered consecutively.
-
-You may use color figures.  However, it is best for the figure captions and the
-paper body to be legible if the paper is printed in either black/white or in
-color.
-
-\subsection{Tables}
-
-All tables must be centered, neat, clean and legible.  The table number and
-title always appear before the table.  See Table~\ref{sample-table}.
-
-Place one line space before the table title, one line space after the
-table title, and one line space after the table. The table title must
-be lower case (except for first word and proper nouns); tables are
-numbered consecutively.
-
-Note that publication-quality tables \emph{do not contain vertical rules.} We
-strongly suggest the use of the \verb+booktabs+ package, which allows for
-typesetting high-quality, professional tables:
-\begin{center}
-  \url{https://www.ctan.org/pkg/booktabs}
-\end{center}
-This package was used to typeset Table~\ref{sample-table}.
-
-\begin{table}
-  \caption{Sample table title}
-  \label{sample-table}
-  \centering
-  \begin{tabular}{lll}
-    \toprule
-    \multicolumn{2}{c}{Part}                   \\
-    \cmidrule(r){1-2}
-    Name     & Description     & Size ($\mu$m) \\
-    \midrule
-    Dendrite & Input terminal  & $\sim$100     \\
-    Axon     & Output terminal & $\sim$10      \\
-    Soma     & Cell body       & up to $10^6$  \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\subsection{Math}
-Note that display math in bare TeX commands will not create correct line numbers for submission.
-Please use LaTeX (or AMSTeX) commands for unnumbered display math.
-(You really shouldn't be using \$\$ anyway; see \url{https://tex.stackexchange.com/questions/503/why-is-preferable-to} and \url{https://tex.stackexchange.com/questions/40492/what-are-the-differences-between-align-equation-and-displaymath} for more information.)
-
-\subsection{Final instructions}
-
-Do not change any aspects of the formatting parameters in the style files.  In
-particular, do not modify the width or length of the rectangle the text should
-fit into, and do not change font sizes (except perhaps in the
-\textbf{References} section; see below). Please note that pages should be
-numbered.
+?
 
@@
 
-Please prepare submission files with paper size ``US Letter,'' and not, for
-example, ``A4.''
+\subsection{Vitals estimation}
 
-Fonts were the main cause of problems in the past years. Your PDF file must only
-contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
-achieve this.
+Table comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens
 
-\begin{itemize}
+\subsection{Which factors impact estimation performance?}
 
-\item You should directly generate PDF files using \verb+pdflatex+.
+Regression analysis to determine factors impacting estimation performance
 
-\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
-  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
-  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
-  available out-of-the-box on most Linux machines.
+\subsection{Impact of subject movement}
 
-\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
-  "solid" shapes instead.
+Bar chart comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens SNR or MAE estimating HR - vs subject movement on x axis.
+Bar chart comparing DeepPhys, MTTS-CAN, VitalLens SNR or MAE estimating RR - vs subject movement on x axis.
 
-\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
-  the equivalent AMS Fonts:
-\begin{verbatim}
-   \usepackage{amsfonts}
-\end{verbatim}
-followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
-for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
-workaround for reals, natural and complex:
-\begin{verbatim}
-   \newcommand{\RR}{I\!\!R} %real numbers
-   \newcommand{\Nat}{I\!\!N} %natural numbers
-   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
-\end{verbatim}
-Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.
+\subsection{Impact of illumination variation}
 
-\end{itemize}
+Bar chart comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens SNR estimating HR - vs illumination variation on x axis.
 
-If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
-you to fix it.
+\subsection{Impact of subject skin type}
 
-\subsection{Margins in \LaTeX{}}
+Bar chart comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens SNR estimating HR - vs subject skin type on x axis.
 
-Most of the margin problems come from figures positioned by hand using
-\verb+\special+ or other commands. We suggest using the command
-\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
-figure width as a multiple of the line width as in the example below:
-\begin{verbatim}
-   \usepackage[pdftex]{graphicx} ...
-   \includegraphics[width=0.8\linewidth]{myfile.pdf}
-\end{verbatim}
-See Section 4.4 in the graphics bundle documentation
-(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})
+\subsection{Impact of subject age}
 
-A number of width problems arise when \LaTeX{} cannot properly hyphenate a
-line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
-necessary.
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2023/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to autmoatically hide this section in the anonymized submission.
-\end{ack}
+Bar chart comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens SNR estimating HR - vs subject age on x axis.
 

### 2023-12-06 paper.tex
Histograms for vitals distributions

@@
 \usepackage{xcolor}         % colors
+\usepackage{pgfplots}
+\usepackage{subcaption}
 
@@
 TODO: Histogram of ages x n participants
+
+\begin{figure}[h!]
+    \centering
+    \begin{subfigure}{0.48\textwidth}
+    	\centering
+    	\begin{tikzpicture}
+        	\begin{axis}[
+            		table/col sep=comma,
+            		ybar,
+            		ylabel=Frequency,
+            		xlabel={Heart rate [bpm]},
+            		xmin=35,
+            		xmax=130,
+            		width=\textwidth,
+            		height=5cm,
+            		grid=major,
+            		bar width=12pt
+            	]
+            	\addplot[fill=red!60,opacity=0.8] table[x index=0, y index=1] {hr_histogram.csv};
+        	\end{axis}
+    	\end{tikzpicture}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\textwidth}
+    	\centering
+    	\begin{tikzpicture}
+        	\begin{axis}[
+            		table/col sep=comma,
+            		ybar,
+            		ylabel=Frequency,
+            		xlabel={Respiratory rate [rpm]},
+            		xmin=-2,
+            		xmax=40,
+            		width=\textwidth,
+            		height=5cm,
+            		grid=major,
+            		bar width=12pt
+            	]
+            	\addplot[fill=blue!60,opacity=0.8] table[x index=0, y index=1] {rr_histogram.csv};
+        	\end{axis}
+    	\end{tikzpicture}
+    \end{subfigure}
+	\begin{subfigure}{0.48\textwidth}
+    	\centering
+    	\begin{tikzpicture}
+        	\begin{axis}[
+            		table/col sep=comma,
+            		ybar,
+            		ylabel=Frequency,
+            		xlabel={Blood pressure [mmHg]},
+            		xmin=50,
+            		xmax=170,
+            		width=\textwidth,
+            		height=5cm,
+            		grid=major,
+            		legend style={at={(.98,.97)}, anchor=north east},
+            	]
+            	\addplot[ybar,bar width=10pt,fill=blue!60,opacity=0.8] table[x index=0, y index=1] {bp_sys_histogram.csv};
+            	\addlegendentry{SBP}
+            	\addplot[ybar,bar width=7pt,fill=red!60,opacity=0.7] table[x index=0, y index=1] {bp_dia_histogram.csv};
+            	\addlegendentry{DBP}
+        	\end{axis}
+    	\end{tikzpicture}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\textwidth}
+    	\centering
+    	\begin{tikzpicture}
+        	\begin{axis}[
+            		table/col sep=comma,
+            		ybar,
+            		ylabel=Frequency,
+            		xlabel={SpO2 [\%]},
+            		xmin=87,
+            		xmax=101,
+            		width=\textwidth,
+            		height=5cm,
+            		grid=major,
+            		bar width=12pt
+            	]
+            	\addplot[fill=blue!60,opacity=0.8] table[x index=0, y index=1] {spo2_histogram.csv};
+        	\end{axis}
+    	\end{tikzpicture}
+    \end{subfigure}
+    \caption{Distributions of chunk summary vitals in PROSIT}
+    \label{fig:histogram}
+\end{figure}
+
 TODO: Histogram of skin types x n participants
-TODO: Histogram of genders x n participants
+TODO: Histogram of genders x n participants1
 

### 2023-12-06 paper.py

@@ -1,0 +1,22 @@
+import numpy as np
+import pandas as pd
+
+export_chunks = pd.read_csv("export_chunks.csv")
+chunks = export_chunks[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type']]
+
+# Histograms
+
+# HR
+frequencies, bin_edges = np.histogram(chunks['frame_avg_hr_pox'], bins=10)
+np.savetxt('hr_histogram.csv', np.column_stack(((bin_edges[:-1]+bin_edges[1:])/2, frequencies)), delimiter=',', header='BinEdges,Frequency', comments='')
+# RR
+frequencies, bin_edges = np.histogram(chunks[(chunks['frame_avg_rr'] >= 0) & (chunks['frame_avg_rr'] <= 40)]['frame_avg_rr'], bins=10)
+np.savetxt('rr_histogram.csv', np.column_stack(((bin_edges[:-1]+bin_edges[1:])/2, frequencies)), delimiter=',', header='BinEdges,Frequency', comments='')
+# BP
+frequencies_sys, bin_edges_sys = np.histogram(chunks[(chunks['frame_avg_bp_sys'] >= 80)]['frame_avg_bp_sys'], bins=10)
+frequencies_dia, bin_edges_dia = np.histogram(chunks[(chunks['frame_avg_bp_dia'] >= 40)]['frame_avg_bp_dia'], bins=10)
+np.savetxt('bp_sys_histogram.csv', np.column_stack(((bin_edges_sys[:-1]+bin_edges_sys[1:])/2, frequencies_sys)), delimiter=',', header='BinEdges,Frequency', comments='')
+np.savetxt('bp_dia_histogram.csv', np.column_stack(((bin_edges_dia[:-1]+bin_edges_dia[1:])/2, frequencies_dia)), delimiter=',', header='BinEdges,Frequency', comments='')
+# SpO2
[…]

### 2023-12-06 paper.tex
Demographics figures

@@
 \usepackage{subcaption}
+\usepackage{csvsimple}
 
@@
 \paragraph{Participant diversity.}
-The skin type diversity of PROSIT is lacking, with too many type 2 and too few type 5 and 6 participants.
-This is why we additionally use the Vital Videos Ghana training set when training our model.
+The demographics of the participants in PROSIT are given in Figure \ref{fig:prosit-demographics}.
+Most participants in PROSIT are under 40 years old, and there are more male than female participants.
+As we show in Section \ref{results}, this is not an issue in practice.
+However, the skin type diversity of PROSIT is also lacking, with too many type 2 and too few type 5 and 6 participants.
+This is in fact an issue.
+To avoid a skin type bias in our model, we supplement our training data with the \textit{Vital Videos Ghana} training set which has mostly type 5 and 6 participants.
 
-TODO: Histogram of ages x n participants
+% Modified example from www.texample.net for pie charts
+% This example needs the packages tikz, xcolor, calc
+\definecolorseries{myseries}{rgb}{step}[rgb]{.95,.85,.55}{.17,.47,.37}
+\resetcolorseries{myseries}%
+
+% a pie slice
+\newcommand{\slice}[4]{
+	\pgfmathsetmacro{\midangle}{0.5*#1+0.5*#2}
+	\begin{scope}
+		\clip (0,0) -- (#1:1) arc (#1:#2:1) -- cycle;
+		\colorlet{SliceColor}{myseries!!+}%
+		\fill[inner color=SliceColor!30,outer color=SliceColor!60] (0,0) circle (1cm);
+	\end{scope}
+	\draw[thick] (0,0) -- (#1:1) arc (#1:#2:1) -- cycle;
+	\node[label={[font=\small]\midangle:#4}] at (\midangle:1) {};
+	\pgfmathsetmacro{\temp}{min((#2-#1-10)/110*(-0.3),0)}
+	\pgfmathsetmacro{\innerpos}{max(\temp,-0.5) + 0.8}
+	\node[font=\small] at (\midangle:\innerpos) {#3};
+}
+
+\begin{figure}[h!]
+	\centering
+    \begin{subfigure}{0.32\textwidth}
+    	\centering
+    	\begin{tikzpicture}
+        	\begin{axis}[
+            		table/col sep=comma,
+            		ybar,
+            		xmin=10,
+            		xmax=95,
+            		ylabel=Frequency,
+            		width=\textwidth,
+            		height=4.6cm,
+            		grid=major,
+            		bar width=8pt,
+            		y label style={yshift=-12pt},
+            		label style={inner sep=0pt}
+            	]
+            	\addplot[fill=blue!60,opacity=0.8] table[x index=0, y index=1] {data/age_histogram.csv};
+        	\end{axis}
+    	\end{tikzpicture}
+    	\caption{Age}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.32\textwidth}
+    	\centering
+    	% drawing of the pie chart
+    	\begin{tikzpicture}[scale=1.5]%
+    	\def\mya{0}\def\myb{0}
+    	\csvreader[before reading=\def\mysum{0}]{data/gender.csv}{Value=\Value}{%
+			\pgfmathsetmacro{\mysum}{\mysum+\Value}%
+		}
+		\csvreader[head to column names]{data/gender.csv}{}{%
+			\let\mya\myb
+			\pgfmathsetmacro{\myb}{\myb+\Value}
+			\slice{\mya/\mysum*360}{\myb/\mysum*360}{\Value}{\Label}
+		}
+		\end{tikzpicture}%
+		\caption{Gender}
+    \end{subfigure}
+    \begin{subfigure}{0.32\textwidth}
+    	\centering
+    	% drawing of the pie chart
+    	\begin{tikzpicture}[scale=1.5]%
+    	% sum of amounts
+		\csvreader[before reading=\def\mysum{0}]{data/gender.csv}{Value=\Value}{%
+			\pgfmathsetmacro{\mysum}{\mysum+\Value}%
+		}
+    	\def\mya{0}\def\myb{0}
+		\csvreader[head to column names]{data/skin_type.csv}{}{%
+			\let\mya\myb
+			\pgfmathsetmacro{\myb}{\myb+\Value}
+			\slice{\mya/\mysum*360}{\myb/\mysum*360}{\Value}{\Label}
+		}
+		\end{tikzpicture}%
+		\caption{Skin type}
+    \end{subfigure}
+    \hfill
+    \caption{Participant demographics in PROSIT}
+    \label{fig:prosit-demographics}
+\end{figure}
+
+\paragraph{Vitals diversity.}
+Distributions of the vitals in PROSIT are given in Figure \ref{fig:prosit-vitals-histogram}.
+The participants are mostly healthy, so these vitals fall in the typical ranges.
+There are several participants who have an irregular heartbeat.
 
@@
             	]
-            	\addplot[fill=red!60,opacity=0.8] table[x index=0, y index=1] {hr_histogram.csv};
+            	\addplot[fill=red!60,opacity=0.8] table[x index=0, y index=1] {data/hr_histogram.csv};
         	\end{axis}
@@
             	]
-            	\addplot[fill=blue!60,opacity=0.8] table[x index=0, y index=1] {rr_histogram.csv};
+            	\addplot[fill=blue!60,opacity=0.8] table[x index=0, y index=1] {data/rr_histogram.csv};
         	\end{axis}
@@
             	]
-            	\addplot[ybar,bar width=10pt,fill=blue!60,opacity=0.8] table[x index=0, y index=1] {bp_sys_histogram.csv};
+            	\addplot[ybar,bar width=10pt,fill=blue!60,opacity=0.8] table[x index=0, y index=1] {data/bp_sys_histogram.csv};
             	\addlegendentry{SBP}
-            	\addplot[ybar,bar width=7pt,fill=red!60,opacity=0.7] table[x index=0, y index=1] {bp_dia_histogram.csv};
+            	\addplot[ybar,bar width=7pt,fill=red!60,opacity=0.7] table[x index=0, y index=1] {data/bp_dia_histogram.csv};
             	\addlegendentry{DBP}
@@
             	]
-            	\addplot[fill=blue!60,opacity=0.8] table[x index=0, y index=1] {spo2_histogram.csv};
+            	\addplot[fill=blue!60,opacity=0.8] table[x index=0, y index=1] {data/spo2_histogram.csv};
         	\end{axis}
@@
     \caption{Distributions of chunk summary vitals in PROSIT}
-    \label{fig:histogram}
+    \label{fig:prosit-vitals-histogram}
 \end{figure}
-
-TODO: Histogram of skin types x n participants
-TODO: Histogram of genders x n participants1
-
-\paragraph{Vitals diversity.}
-There are also several participants who have an irregular heartbeat.
-
-TODO: Histogram of HR x n chunks
-TODO: Histogram of RR x n chunks
-TODO: Histogram of BP x n chunks
-TODO: Histogram of SpO2 x n chunks
-
 \subsection{Vital Videos}
@@
 \section{Results and Discussion}
+\label{results}
 

### 2023-12-06 paper.py

@@ -3,20 +3,37 @@ import pandas as pd
 
 export_chunks = pd.read_csv("export_chunks.csv")
 chunks = export_chunks[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type']]
+sessions = chunks.drop_duplicates(subset='session_id')
 
-# Histograms
+# Age
+frequencies, bin_edges = np.histogram(sessions['subject_age'], bins=8)
+np.savetxt('data/age_histogram.csv', np.column_stack(((bin_edges[:-1]+bin_edges[1:])/2, frequencies)), delimiter=',', header='BinEdges,Frequency', comments='')
+
+# Gender
+counts = sessions['subject_gender'].value_counts()
+counts_df = pd.DataFrame({'Label': counts.index.str.capitalize(), 'Value': counts.values})
+counts_df.to_csv("data/gender.csv", index=False)
+
+# Skin Type
+counts = sessions['subject_skin_type'].value_counts()
+counts_df = pd.DataFrame({'Label': counts.index.map(lambda val: "Type {}".format(val)), 'Value': counts.values})
+counts_df = counts_df.sort_values(by='Label')
+counts_df.to_csv("data/skin_type.csv", index=False)
 
 # HR
 frequencies, bin_edges = np.histogram(chunks['frame_avg_hr_pox'], bins=10)
-np.savetxt('hr_histogram.csv', np.column_stack(((bin_edges[:-1]+bin_edges[1:])/2, frequencies)), delimiter=',', header='BinEdges,Frequency', comments='')
+np.savetxt('data/hr_histogram.csv', np.column_stack(((bin_edges[:-1]+bin_edges[1:])/2, frequencies)), delimiter=',', header='BinEdges,Frequency', comments='')
+
 # RR
[…]

### 2023-12-07 paper.tex
Automatically draw tables from csv; Add bibliography; Re-structure dataset section

@@
 \usepackage{csvsimple}
+\usepackage{siunitx}
+\sisetup{group-separator = {,}, group-minimum-digits = 4}
 
-\title{VitalLens 1.0 Technical Report}
+\title{VitalLens: Take A Vital Selfie}
 
@@
 \begin{abstract}
-We report the development of VitalLens, an app that estimates vital signs such as heart rate and respiration rate from selfie video in real time.
+This report introduces VitalLens, an app that estimates vital signs such as heart rate and respiration rate from selfie video in real time.
 The estimation engine of VitalLens is a computer vision model trained on a diverse dataset of video and gold-standard ground truth physiological sensor data.
@@
 
-Video of the human face and upper body contains signals with information about the subject's physiological state.
+Video of the human face and upper body contains signals with information about the subject's physiological state \citep{verkruysse2008remote}.
 Given appropriate circumstances, these signals can be strong enough to estimate vital signs such as heart rate and respiratory rate.
@@
 
-Paragraph on G, CHROM, and POS.
+Paragraph on G \cite{verkruysse2008remote}, CHROM \cite{de2013robust}, and POS \cite{wang2017algorithmic}.
 
-Paragraph on DeepPhys and MTTS-CAN.
+Paragraph on DeepPhys \cite{chen2018deep} and MTTS-CAN \cite{liu2020multi}.
 
 Acknowledge other models we don't compare because lack focus on real-time.
+
+[INSERT FIGURE WITH SAMPLE PREDICTIONS FROM VALIDATION SET]
+
+Section summarizing the remainder of the report.
 
@@
 This report focuses on the capabilities and limitations of VitalLens.
-The estimation engine of VitalLens is a recurrent convolutional model trained on a diverse dataset of video and gold-standard ground truth physiological data.
+The estimation engine of VitalLens is a computer vision model trained on a diverse dataset of video and gold-standard ground truth physiological data.
 
@@
 
-Two major datasets are involved with training and evaluation of VitalLens:
-The \textit{PROSIT} dataset and the \textit{Vital Videos} dataset.
-The vast majority of the data used to train VitalLens comes from PROSIT.
-We enrich the training data with a subset of data from the Vital Videos dataset, as detailed in this section.
-The main Vital Videos dataset serves as a test set for the evaluation of VitalLens.
+VitalLens is trained on the \textit{PROSIT} and \textit{Vital Videos Ghana} datasets.
+The vast majority of training data comes from PROSIT.
+For evaluation, we use the entire \textit{Vital Videos Medium} dataset as well as the test sets of the \textit{PROSIT} and \textit{Vital Videos Ghana} datasets.
 
@@
 
-PROSIT (\underline{P}hysiological \underline{R}ecordings \underline{O}f \underline{S}ubjects using \underline{I}maging \underline{T}echnology) is our in-house dataset for practical rPPG applications.
+PROSIT (\underline{P}hysiological \underline{R}ecordings \underline{O}f \underline{S}ubjects using \underline{I}maging \underline{T}echnology) is our in-house dataset collected in Australia for practical rPPG applications.
 
@@
 \paragraph{Dataset size and split.}
+\csvreader[before reading=\def\totalparticipantsprosit{0}\def\totalchunksprosit{0}\def\totaltimeprosit{0}\def\totalsessionsprosit{0}]{data/prosit_summary.csv}{participants=\participants,chunks=\chunks,time=\time,sessions=\sessions}{%
+	\pgfmathsetmacro{\totalparticipantsprosit}{\totalparticipantsprosit+\participants}%
+	\pgfmathsetmacro{\totalchunksprosit}{\totalchunksprosit+\chunks}%
+	\pgfmathsetmacro{\totaltimeprosit}{\totaltimeprosit+\time}%
+	\pgfmathsetmacro{\totalsessionsprosit}{\totalsessionsprosit+\sessions}%
+}
 Development of PROSIT is ongoing.
-As of the writing of this report, it comprises 157 unique participants across 173 recording sessions in 45 different locations.
-This results in a total of 9,767 chunks or 27.8 hours of data.
+As of the writing of this report, it comprises \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsprosit} unique participants across \sisetup{round-mode=places,round-precision=0}\num{\totalsessionsprosit} recording sessions in 45 different locations.
+This results in a total of \sisetup{round-mode=places,round-precision=0}\num{\totalchunksprosit} chunks or \sisetup{round-mode=places,round-precision=1}\num{\totaltimeprosit} hours of data.
 
@@
  	\caption{PROSIT Dataset Size}
-  \label{prosit-table}
+ 	\label{tab:prosit-summary}
  	\centering
-  \begin{tabular}{llll}
-    \toprule
-    Split      & \# Participants & \# Chunks & Time   \\
-    \midrule
-    Training   & 114 			 & 6,765     & 19.4 h \\
-    Validation & 23 			 & 1,599     & 4.5 h  \\
-    Test       & 20     			 & 1,403     & 3.9 h  \\
-    \midrule
-    Total      & 157			 & 9,767  	 & 27.8 h \\
-    \bottomrule
-  \end{tabular}
+  	\csvreader[
+		tabular=llll,
+    	table head=\toprule Split & \# Participants & \# Chunks & Time \\ \midrule,
+    	table foot=\midrule Total & \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsprosit} & \sisetup{round-mode=places,round-precision=0}\num{\totalchunksprosit} & \sisetup{round-mode=places,round-precision=1}\SI{\totaltimeprosit}{\hour} \\ \bottomrule,
+    	late after line=\\,
+    	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
+  	]{data/prosit_summary.csv}{}%
+  {\csvcoli & \num{\csvcolii} & \num{\csvcoliv} & \SI{\csvcolv}{\hour} }
 \end{table}
@@
 
-\paragraph{Participant diversity.}
-The demographics of the participants in PROSIT are given in Figure \ref{fig:prosit-demographics}.
-Most participants in PROSIT are under 40 years old, and there are more male than female participants.
-As we show in Section \ref{results}, this is not an issue in practice.
-However, the skin type diversity of PROSIT is also lacking, with too many type 2 and too few type 5 and 6 participants.
-This is in fact an issue.
-To avoid a skin type bias in our model, we supplement our training data with the \textit{Vital Videos Ghana} training set which has mostly type 5 and 6 participants.
+\subsection{Vital Videos Medium}
+
+Vital Videos is a large, diverse dataset for rPPG collected in Belgium \cite{toye2023vital}.
+It is the largest dataset we are aware of that is available for research without academic affiliation.
+We use a slightly extended version of the medium instantiation (``VV-Medium''), which consists of 296 participants.
+
+\paragraph{Pre-processing.}
+
+We pre-process VV-Medium using the same steps applied to PROSIT.
+As part of this step, we create small chunks, calculate summary vitals, and extract further metadata.
+
+\begin{table}[h!]
+ 	\caption{VV-Medium Dataset Size}
+ 	\label{tab:vv-medium-summary}
+ 	\centering
+  	\csvreader[
+		tabular=llll,
+    	table head=\toprule Split & \# Participants & \# Chunks & Time \\ \midrule,
+    	table foot=\bottomrule,
+    	late after line=\\,
+    	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
+  	]{data/vv_medium_summary.csv}{}%
+  {\csvcoli & \num{\csvcolii} & \num{\csvcoliii} & \SI{\csvcoliv}{\hour} }
+\end{table}
+
+The entirety of VV-Medium is used to test the capabilities of VitalLens.
+
+\subsection{Vital Videos Ghana}
+
+Vital Videos Ghana is a new dataset from the authors of Vital Videos aiming to address the insufficient share of participants with skin types 5 and 6 usually found in rPPG datasets.
+It was collected in Ghana, with the majority of participants having skin type 5 or 6.
+\csvreader[before reading=\def\totalparticipantsvvghana{0}\def\totalchunksvvghana{0}\def\totaltimevvghana{0}]{data/vv_ghana_small_summary.csv}{participants=\participants,chunks=\chunks,time=\time}{%
+	\pgfmathsetmacro{\totalparticipantsvvghana}{\totalparticipantsvvghana+\participants}%
+	\pgfmathsetmacro{\totalchunksvvghana}{\totalchunksvvghana+\chunks}%
+	\pgfmathsetmacro{\totaltimevvghana}{\totaltimevvghana+\time}%
+}
+We use a small instantiation (``VV-Ghana-Small''), which consists of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvghana} participants.
+
+\paragraph{Pre-processing.}
+
+We pre-process VV-Ghana-Small using the same steps applied to PROSIT.
+As part of this step, we create small chunks, calculate summary vitals, and extract further metadata.
+
+\begin{table}[h!]
+ 	\caption{VV-Ghana-Small Dataset Size}
+ 	\label{tab:vv-ghana-small-summary}
+ 	\centering
+  	\csvreader[
+		tabular=llll,
+    	table head=\toprule Split & \# Participants & \# Chunks & Time \\ \midrule,
+    	table foot=\midrule Total & \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvghana} & \sisetup{round-mode=places,round-precision=0}\num{\totalchunksvvghana} & \sisetup{round-mode=places,round-precision=1}\SI{\totaltimevvghana}{\hour} \\ \bottomrule,
+    	late after line=\\,
+    	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
+  	]{data/vv_ghana_small_summary.csv}{}%
+  {\csvcoli & \num{\csvcolii} & \num{\csvcoliii} & \sisetup{round-mode=places,round-precision=1}\SI{\csvcoliv}{\hour} }
+\end{table}
+
+Each participant is randomly assigned to be part of either the \textit{training}, \textit{validation}, or \textit{test} set to ensure that all participants seen during validation and test are previously unseen by the model.
+
+\subsection{Final training dataset: PROSIT + VV-Ghana-Small}
+\csvreader[before reading=\def\totalparticipantstraining{0}\def\totalchunkstraining{0}\def\totaltimetraining{0}]{data/training_summary.csv}{participants=\participants,chunks=\chunks,time=\time}{%
+		\pgfmathsetmacro{\totalparticipantstraining}{\totalparticipantstraining+\participants}%
+		\pgfmathsetmacro{\totalchunkstraining}{\totalchunkstraining+\chunks}%
+		\pgfmathsetmacro{\totaltimetraining}{\totaltimetraining+\time}%
+	}
+We combine the training sets of PROSIT and VV-Ghana-Small for training.
+As of this writing, this makes a total of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantstraining} unique participants, \sisetup{round-mode=places,round-precision=0}\num{\totalchunkstraining} chunks or \sisetup{round-mode=places,round-precision=1}\num{\totaltimetraining} hours of data.
+
+\begin{table}[h!]
+ 	\caption{VitalLens Training Dataset Size}
+ 	\label{tab:training-summary}
+ 	\centering
+  	\csvreader[
+		tabular=llll,
+    	table head=\toprule Source & \# Participants & \# Chunks & Time \\ \midrule,
+    	table foot=\midrule Total & \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantstraining} & \sisetup{round-mode=places,round-precision=0}\num{\totalchunkstraining} & \sisetup{round-mode=places,round-precision=1}\SI{\totaltimetraining}{\hour} \\ \bottomrule,
+    	late after line=\\,
+    	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
+  	]{data/training_summary.csv}{}%
+  {\csvcoli & \num{\csvcolii} & \num{\csvcoliii} & \sisetup{round-mode=places,round-precision=1}\SI{\csvcoliv}{\hour} }
+\end{table}
+
+\paragraph{Dataset demographics.}
+The demographics of our training dataset are given in Figure \ref{fig:training-demographics}.
+The participants are predominantly on the younger side, but as we show in Section \ref{results}, this is not an issue in practice.
+Genders are equally represented.
+Although the skin type diversity of PROSIT by itself is lacking, this combined training dataset has a diverse representation of skin types.
+As we show in Section \ref{results}, this helps us to reduce the skin type bias of VitalLens.
 
@@
     	% sum of amounts
-		\csvreader[before reading=\def\mysum{0}]{data/gender.csv}{Value=\Value}{%
+		\csvreader[before reading=\def\mysum{0}]{data/skin_type.csv}{Value=\Value}{%
 			\pgfmathsetmacro{\mysum}{\mysum+\Value}%
@@
     \hfill
-    \caption{Participant demographics in PROSIT}
-    \label{fig:prosit-demographics}
+    \caption{Participant demographics in training dataset}
+    \label{fig:training-demographics}
 \end{figure}
@@
 \paragraph{Vitals diversity.}
-Distributions of the vitals in PROSIT are given in Figure \ref{fig:prosit-vitals-histogram}.
+Distributions of the vitals in our training dataset are given in Figure \ref{fig:training-vitals-histogram}.
 The participants are mostly healthy, so these vitals fall in the typical ranges.
@@
     \end{subfigure}
-	\begin{subfigure}{0.48\textwidth}
-    	\centering
-    	\begin{tikzpicture}
-        	\begin{axis}[
-            		table/col sep=comma,
-            		ybar,
-            		ylabel=Frequency,
-            		xlabel={Blood pressure [mmHg]},
-            		xmin=50,
-            		xmax=170,
-            		width=\textwidth,
-            		height=5cm,
-            		grid=major,
-            		legend style={at={(.98,.97)}, anchor=north east},
-            	]
-            	\addplot[ybar,bar width=10pt,fill=blue!60,opacity=0.8] table[x index=0, y index=1] {data/bp_sys_histogram.csv};
-            	\addlegendentry{SBP}
-            	\addplot[ybar,bar width=7pt,fill=red!60,opacity=0.7] table[x index=0, y index=1] {data/bp_dia_histogram.csv};
-            	\addlegendentry{DBP}
-        	\end{axis}
-    	\end{tikzpicture}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.48\textwidth}
-    	\centering
-    	\begin{tikzpicture}
-        	\begin{axis}[
-            		table/col sep=comma,
-            		ybar,
-            		ylabel=Frequency,
-            		xlabel={SpO2 [\%]},
-            		xmin=87,
-            		xmax=101,
-            		width=\textwidth,
-            		height=5cm,
-            		grid=major,
-            		bar width=12pt
-            	]
-            	\addplot[fill=blue!60,opacity=0.8] table[x index=0, y index=1] {data/spo2_histogram.csv};
-        	\end{axis}
-    	\end{tikzpicture}
-    \end{subfigure}
-    \caption{Distributions of chunk summary vitals in PROSIT}
-    \label{fig:prosit-vitals-histogram}
+    \caption{Distributions of chunk summary vitals in training dataset}
+    \label{fig:training-vitals-histogram}
 \end{figure}
-\subsection{Vital Videos}
-
-\subsubsection{Vital Videos Medium}
-
-Used for test.
-
-Pieter's dataset
-Number of unique subjects
-Total duration
-Statistics
-Distributions of metadata
-
-\subsubsection{Vital Videos Ghana}
-
-Used for training.
-
-Pieter's dataset
-Number of unique subjects
-Total duration
-Statistics
-Distributions of metadata
 
@@
 
-References follow the acknowledgments in the camera-ready paper. Use unnumbered first-level heading for
-the references. Any choice of citation style is acceptable as long as you are
-consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
-when listing the references.
-Note that the Reference section does not count towards the page limit.
-\medskip
-
-{
-\small
-
-[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
-connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
-(eds.), {\it Advances in Neural Information Processing Systems 7},
-pp.\ 609--616. Cambridge, MA: MIT Press.
-
-[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
-  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
-TELOS/Springer--Verlag.
-
-[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
-recall at excitatory recurrent synapses and cholinergic modulation in rat
-hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
-}
+\bibliographystyle{plain}
+\bibliography{paper}
 

### 2023-12-07 paper.py

@@ -1,39 +1,108 @@
 import numpy as np
 import pandas as pd
 
-export_chunks = pd.read_csv("export_chunks.csv")
-chunks = export_chunks[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type']]
-sessions = chunks.drop_duplicates(subset='session_id')
+# == Import data from csv ==
 
-# Age
-frequencies, bin_edges = np.histogram(sessions['subject_age'], bins=8)
+# PROSIT
+chunks_prosit = pd.read_csv("export_prosit.csv")[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type']]
+chunks_prosit['subject_id'] = chunks_prosit['session_description'].str.split('-').str.get(0)
+sessions_prosit = chunks_prosit.drop_duplicates(subset='session_id')
+# VV
+chunks_vv = pd.read_csv("export_vv_medium.csv")[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type']]
[…]

### 2023-12-07 paper.tex
Methodology section

@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 
-\section{Scope and Limitations of this Technical Report}
-\label{gen_inst}
+\section{Model Architecture}
+\label{sec:scope}
 
-This report focuses on the capabilities and limitations of VitalLens.
-The estimation engine of VitalLens is a computer vision model trained on a diverse dataset of video and gold-standard ground truth physiological data.
+The estimation engine built into VitalLens is a computer vision model trained on a diverse dataset of video and gold-standard ground truth physiological data.
+The model architecture is broadly based on the \textit{EfficientNetV2} \cite{tan2021efficient} model family, enhanced with several improvements in architecture and model optimization to enable efficient training and inference in the rPPG domain.
 
-This report does not contain any further detail about the architecture of the model or training methodology -
-we reserve this for future publication.
+This report does not contain any further detail about the architecture of the model or training methodology - we reserve this for future publication.
 
 \section{Datasets}
-\label{headings}
+\label{sec:datasets}
 
@@
 Each potential participant was required to go through an informed consent process in accordance with Australian privacy laws prior to participation.
-During the session, participants are asked to complete tasks on an iPad while sensor data is being recorded.
+During the session, participants are asked to complete tasks on a hand-held iPad while sensor data is being recorded.
+Participants are not explicitly asked to remain stationary, which results in varying amounts of participant movement.
+For some sessions, the camera is on a tripod and thereby stationary, while for other sessions the camera is fixed to the iPad and thereby not necessarily stationary.
 
@@
 We use a slightly extended version of the medium instantiation (``VV-Medium''), which consists of 296 participants.
+Both camera and participants are stationary in this dataset.
 
@@
 We use a small instantiation (``VV-Ghana-Small''), which consists of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvghana} participants.
+Both camera and participants are stationary in this dataset.
 
@@
 The demographics of our training dataset are given in Figure \ref{fig:training-demographics}.
-The participants are predominantly on the younger side, but as we show in Section \ref{results}, this is not an issue in practice.
+The participants are predominantly on the younger side, but as we show in Section \ref{sec:results}, this is not an issue in practice.
 Genders are equally represented.
 Although the skin type diversity of PROSIT by itself is lacking, this combined training dataset has a diverse representation of skin types.
-As we show in Section \ref{results}, this helps us to reduce the skin type bias of VitalLens.
+As we show in Section \ref{sec:results}, this helps us to reduce the skin type bias of VitalLens.
 
@@
 
-\paragraph{Vitals diversity.}
+\paragraph{Dataset vitals diversity.}
 Distributions of the vitals in our training dataset are given in Figure \ref{fig:training-vitals-histogram}.
@@
 \section{Methodology}
-\label{others}
+\label{sec:methodology}
 
-?
+To develop and evaluate the VitalLens estimation engine model, we follow a systematic approach involving model training, validation, and testing:
+
+\paragraph{Model Training.}
+The training process optimizes the model parameters to learn the mapping between facial video signals and ground truth physiological data.
+For this purpose, we use a diverse training dataset of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantstraining} unique participants introduced in Section \ref{sec:datasets}.
+
+\paragraph{Model Validation.}
+While developing the model, we train many different versions of our model.
+During training, we continuously monitor the model's performance on the validation sets of PROSIT and VV-Ghana-Small as outlined in Section \ref{sec:datasets}.
+Since the training dataset is disjoint from both validation sets in terms of participants, this allows us to calculate validation metrics which measure how well each model has learned to generalize in estimating vital signs from video.
+We used the validation metrics to select the final model to be used in VitalLens.
+
+\paragraph{Model Testing.}
+The final evaluation of the VitalLens estimation engine is conducted by measuring the generalization abilities of the final model on several datasets, all of which are participant-disjoint from both the training and validation datasets:
+
+\begin{itemize}
+	\item The entire VV-Medium dataset, a large and diverse dataset of 289 participants introduced in Section \ref{sec:datasets}.
+It includes participants with varying demographics and environmental conditions.
+	\item The test set of the PROSIT dataset. This is important as PROSIT includes camera and participant movement, which is to be expected when VitalLens is used in the wild.
+	\item The test set of the VV-Ghana-Small dataset.
+\end{itemize}
+
+The evaluation metrics include key vital signs such as heart rate and respiratory rate.
+We compare the performance of VitalLens with existing methods, including G, CHROM, POS, DeepPhys, and MTTS-CAN, using metrics such as Signal-to-Noise Ratio (SNR) and Mean Absolute Error (MAE).
+
+This methodology ensures a thorough assessment of the VitalLens estimation engine, considering diverse data sources and real-world conditions.
+The results and discussions in the following sections provide insights into the effectiveness and limitations of the proposed approach.
 
 \section{Results and Discussion}
-\label{results}
+\label{sec:results}
 
@@
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2023-12-07 paper.tex
Work on model evaluation

@@
 We use a slightly extended version of the medium instantiation (``VV-Medium''), which consists of 296 participants.
+% TODO pull from data
 Both camera and participants are stationary in this dataset.
@@
 
-The evaluation metrics include key vital signs such as heart rate and respiratory rate.
+The evaluation metrics focus on the key vital signs, \textit{heart rate} and \textit{respiratory rate}.
 We compare the performance of VitalLens with existing methods, including G, CHROM, POS, DeepPhys, and MTTS-CAN, using metrics such as Signal-to-Noise Ratio (SNR) and Mean Absolute Error (MAE).
+
+We furthermore conduct a regression analysis using the chunk metadata to predict the performance of VitalLens.
+This helps us identify the environmental factor affecting performance, and informs ``how to use'' advice offered to users of VitalLens.
 
@@
 
-Table comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens
+TODO evaluate for all methods
+
+\begin{table}[h!]
+ 	\caption{Vitals estimation results on VV-Medium}
+ 	\label{tab:results-vv-medium}
+ 	\centering
+  	\csvreader[
+		tabular=lcccccc,
+    	table head=\toprule Method & \multicolumn{1}{c}{Heart rate} & \multicolumn{2}{c}{Pulse wave} & \multicolumn{1}{c}{Respiratory rate} & \multicolumn{2}{c}{Respiration wave} \\ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ \\ \midrule,
+    	table foot=\bottomrule,
+    	late after line=\\,
+    	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
+  	]{data/results_vv_medium.csv}{}%
+  {\csvcoli & \sisetup{round-mode=places,round-precision=2}\num{\csvcolii} & \sisetup{round-mode=places,round-precision=2}\num{\csvcoliii} & \sisetup{round-mode=places,round-precision=2}\num{\csvcoliv} & \sisetup{round-mode=places,round-precision=2}\num{\csvcolv} & \sisetup{round-mode=places,round-precision=2}\num{\csvcolvi} & \sisetup{round-mode=places,round-precision=2}\num{\csvcolvii} }
+\end{table}
+
+\begin{table}[h!]
+ 	\caption{Vitals estimation results on PROSIT test set}
+ 	\label{tab:results-prosit-test}
+ 	\centering
+  	\csvreader[
+		tabular=lcccccc,
+    	table head=\toprule Method & \multicolumn{1}{c}{Heart rate} & \multicolumn{2}{c}{Pulse wave} & \multicolumn{1}{c}{Respiratory rate} & \multicolumn{2}{c}{Respiration wave} \\ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ \\ \midrule,
+    	table foot=\bottomrule,
+    	late after line=\\,
+    	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
+  	]{data/results_prosit_test.csv}{}%
+  {\csvcoli & \sisetup{round-mode=places,round-precision=2}\num{\csvcolii} & \sisetup{round-mode=places,round-precision=2}\num{\csvcoliii} & \sisetup{round-mode=places,round-precision=2}\num{\csvcoliv} & \sisetup{round-mode=places,round-precision=2}\num{\csvcolv} & \sisetup{round-mode=places,round-precision=2}\num{\csvcolvi} & \sisetup{round-mode=places,round-precision=2}\num{\csvcolvii} }
+\end{table}
+
+\begin{table}[h!]
+ 	\caption{Vitals estimation results on VV-Ghana-Small test set}
+ 	\label{tab:results-vv-ghana-small-test}
+ 	\centering
+  	\csvreader[
+		tabular=lcccccc,
+    	table head=\toprule Method & \multicolumn{1}{c}{Heart rate} & \multicolumn{2}{c}{Pulse wave} & \multicolumn{1}{c}{Respiratory rate} & \multicolumn{2}{c}{Respiration wave} \\ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ \\ \midrule,
+    	table foot=\bottomrule,
+    	late after line=\\,
+    	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
+  	]{data/results_vv_ghana_test.csv}{}%
+  {\csvcoli & \sisetup{round-mode=places,round-precision=2}\num{\csvcolii} & \sisetup{round-mode=places,round-precision=2}\num{\csvcoliii} & \sisetup{round-mode=places,round-precision=2}\num{\csvcoliv} & \sisetup{round-mode=places,round-precision=2}\num{\csvcolv} & \sisetup{round-mode=places,round-precision=2}\num{\csvcolvi} & \sisetup{round-mode=places,round-precision=2}\num{\csvcolvii} }
+\end{table}
 

### 2023-12-07 paper.py

@@ -4,15 +4,15 @@ import pandas as pd
 # == Import data from csv ==
 
 # PROSIT
-chunks_prosit = pd.read_csv("export_prosit.csv")[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type']]
+chunks_prosit = pd.read_csv("export_prosit.csv")[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'env_illuminance', 'env_illuminance_d', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate']]
 chunks_prosit['subject_id'] = chunks_prosit['session_description'].str.split('-').str.get(0)
 sessions_prosit = chunks_prosit.drop_duplicates(subset='session_id')
 # VV
-chunks_vv = pd.read_csv("export_vv_medium.csv")[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type']]
[…]

### 2023-12-08 paper.tex
Work on evaluation

--- 2023-12-07 paper.tex
@@
 \sisetup{group-separator = {,}, group-minimum-digits = 4}
+\usepackage[flushleft]{threeparttable} % For tablenotes
 
@@
 
-Video of the human face and upper body contains signals with information about the subject's physiological state \citep{verkruysse2008remote}.
-Given appropriate circumstances, these signals can be strong enough to estimate vital signs such as heart rate and respiratory rate.
-This process is known as Remote Photoplethysmography (rPPG); many rPPG approaches have been proposed, including handcrafted algorithms and ones learned from empirical data.
-Depending on the intended application, we can measure advantages and drawbacks of rPPG approaches in the accuracy, inference speed, and robustness regarding various factors impacting estimation performance.
+Video of the human face and upper body has proven to be a rich source of information about an individual's physiological state \cite{verkruysse2008remote}.
+In particular, signals embedded in these videos can be harnessed to estimate vital signs such as heart rate and respiratory rate through a process known as Remote Photoplethysmography (rPPG).
+This capability holds immense potential for non-invasive and real-time health monitoring applications.
 
-Paragraph on G \cite{verkruysse2008remote}, CHROM \cite{de2013robust}, and POS \cite{wang2017algorithmic}.
+Various rPPG approaches have been proposed, including handcrafted algorithms and ones learned from empirical data.
+Handcrafted algorithms offer advantages of fast inference speed and no training data required, but usually lack in accuracy; notable approaches include the original approach G \cite{verkruysse2008remote}, and the more sophisticated CHROM \cite{de2013robust} and POS \cite{wang2017algorithmic}.
+Given enough high-quality data to learn from, learning-based approaches such as DeepPhys \cite{chen2018deep} and MTTS-CAN \cite{liu2020multi} hold the promise of greater accuracy, which they have to trade off against inference speed.
+There are also other learning-based approaches achieving high accuracy, which we do not cover here as they are not aiming to achieve real-time inference.
 
-Paragraph on DeepPhys \cite{chen2018deep} and MTTS-CAN \cite{liu2020multi}.
-
-Acknowledge other models we don't compare because lack focus on real-time.
+In this context, we introduce VitalLens, an innovative application designed to estimate vital signs, including heart rate and respiratory rate, in real time from selfie videos.
+The underlying estimation engine of VitalLens is built upon a computer vision model trained on a diverse dataset, combining video recordings with gold-standard ground truth physiological sensor data.
 
@@
 
-Section summarizing the remainder of the report.
+This report focuses on the capabilities and limitation of VitalLens.
+To this end, We present a comprehensive performance evaluation, comparing it with the named existing methods on diverse datasets.
+
+The remainder of this report is organized as follows: Section \ref{sec:scope} provides an brief overview of the model, Section \ref{sec:datasets} details the datasets used for training and evaluation, Section \ref{sec:methodology} outlines the systematic approach to model development and evaluation, Section \ref{sec:results} presents the results and discussions, and finally, Section \ref{sec:conclusion} concludes the report with implications and future directions.
 
@@
 This yields a rich set of data including video, ECG, PPG, SpO2, respiration, blood pressure, and ambient luminance.
-We also collect age, gender, height, and skin type metadata according to the Fitzpatrick scale.
+We also collect age, gender, height, and skin type metadata according to the Fitzpatrick scale \cite{fitzpatrick1975soleil}.
 
@@
 It is the largest dataset we are aware of that is available for research without academic affiliation.
-We use a slightly extended version of the medium instantiation (``VV-Medium''), which consists of 296 participants.
+\csvreader[before reading=\def\totalparticipantsvvmedium{0}\def\totalchunksvvmedium{0}\def\totaltimevvmedium{0}]{data/vv_medium_summary.csv}{participants=\participants,chunks=\chunks,time=\time}{%
+	\pgfmathsetmacro{\totalparticipantsvvmedium}{\totalparticipantsvvmedium+\participants}%
+	\pgfmathsetmacro{\totalchunksvvmedium}{\totalchunksvvmedium+\chunks}%
+	\pgfmathsetmacro{\totaltimevvmedium}{\totaltimevvmedium+\time}%
+}
+We use a slightly extended version of the medium instantiation (``VV-Medium''), which consists of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvmedium} participants.
 Both camera and participants are stationary in this dataset.
@@
 As part of this step, we create small chunks, calculate summary vitals, and extract further metadata.
+Note that for many chunks, the missing respiratory signal was synthetically created with an older version of VitalLens and then manually verified.
 
@@
   	]{data/vv_medium_summary.csv}{}%
-  {\csvcoli & \num{\csvcolii} & \num{\csvcoliii} & \SI{\csvcoliv}{\hour} }
+  {\csvcoli & \num{\csvcolii} & \num{\csvcoliii} & \sisetup{round-mode=places,round-precision=1}\SI{\csvcoliv}{\hour} }
 \end{table}
@@
 As part of this step, we create small chunks, calculate summary vitals, and extract further metadata.
+Note that for some chunks, the missing respiratory signal was synthetically created with an older version of VitalLens and then manually verified.
 
@@
 
-% Modified example from www.texample.net for pie charts
-% This example needs the packages tikz, xcolor, calc
-\definecolorseries{myseries}{rgb}{step}[rgb]{.95,.85,.55}{.17,.47,.37}
-\resetcolorseries{myseries}%
+% Define color series for the first pie chart
+\definecolorseries{myseries1}{rgb}{step}[rgb]{1,0,0}{-1,0,1}
+\resetcolorseries{myseries1}%
+
+% Define color series for the second pie chart (six shades of gray)
+\definecolorseries{myseries2}{rgb}{step}[rgb]{1.0, 0.75, 0.65}{-0.134,-0.122,-0.13}
+\resetcolorseries{myseries2}%
 
 % a pie slice
-\newcommand{\slice}[4]{
+\newcommand{\slice}[5]{
 	\pgfmathsetmacro{\midangle}{0.5*#1+0.5*#2}
@@
 		\clip (0,0) -- (#1:1) arc (#1:#2:1) -- cycle;
-		\colorlet{SliceColor}{myseries!!+}%
+		\colorlet{SliceColor}{#3!!+}%%
 		\fill[inner color=SliceColor!30,outer color=SliceColor!60] (0,0) circle (1cm);
@@
 	\draw[thick] (0,0) -- (#1:1) arc (#1:#2:1) -- cycle;
-	\node[label={[font=\small]\midangle:#4}] at (\midangle:1) {};
+	\node[label={[font=\small]\midangle:#5}] at (\midangle:1) {};
 	\pgfmathsetmacro{\temp}{min((#2-#1-10)/110*(-0.3),0)}
 	\pgfmathsetmacro{\innerpos}{max(\temp,-0.5) + 0.8}
-	\node[font=\small] at (\midangle:\innerpos) {#3};
+	\node[font=\small] at (\midangle:\innerpos) {#4};
 }
@@
 			\pgfmathsetmacro{\myb}{\myb+\Value}
-			\slice{\mya/\mysum*360}{\myb/\mysum*360}{\Value}{\Label}
+			\slice{\mya/\mysum*360}{\myb/\mysum*360}{myseries1}{\Value}{\Label}
 		}
@@
 			\pgfmathsetmacro{\myb}{\myb+\Value}
-			\slice{\mya/\mysum*360}{\myb/\mysum*360}{\Value}{\Label}
+			\slice{\mya/\mysum*360}{\myb/\mysum*360}{myseries2}{\Value}{\Label}
 		}
@@
 \paragraph{Model Testing.}
-The final evaluation of the VitalLens estimation engine is conducted by measuring the generalization abilities of the final model on several datasets, all of which are participant-disjoint from both the training and validation datasets:
+The final evaluation of the VitalLens estimation engine is conducted by benchmarking the generalization abilities of the final model on several datasets, all of which are participant-disjoint from both the training and validation datasets:
 
 \begin{itemize}
-	\item The entire VV-Medium dataset, a large and diverse dataset of 289 participants introduced in Section \ref{sec:datasets}.
+	\item The entire VV-Medium dataset, a large and diverse dataset of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvmedium} participants introduced in Section \ref{sec:datasets}.
 It includes participants with varying demographics and environmental conditions.
@@
 
-The evaluation metrics include key vital signs such as heart rate and respiratory rate.
-We compare the performance of VitalLens with existing methods, including G, CHROM, POS, DeepPhys, and MTTS-CAN, using metrics such as Signal-to-Noise Ratio (SNR) and Mean Absolute Error (MAE).
-
-This methodology ensures a thorough assessment of the VitalLens estimation engine, considering diverse data sources and real-world conditions.
-The results and discussions in the following sections provide insights into the effectiveness and limitations of the proposed approach.
-
 \section{Results and Discussion}
@@
 
+We evaluated VitalLens on three datasets - VV-Medium, the PROSIT test set, and the VV-Ghana-Small test set.
+None of the participants included in these datasets were previously seen by VitalLens.
+In addition, we also evaluated several other existing methods to allow benchmarking of VitalLens:
+The handcrafted methods G, CHROM, and POS, as well as the learning-based methods DeepPhys and MTTS-CAN.
+To allow a fair comparison, we trained the latter two on the same training data as VitalLens.
+
+Furthermore, we conduct a regression analysis by using metadata to predict the performance of VitalLens.
+This helps us identify the factors affecting performance, and informs ``how to use'' advice offered to users of VitalLens.
+
+Finally, we analyze the influence of the factors of interest in more detail.
+
 \subsection{Vitals estimation}
 
-Table comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens
+Vitals estimation is performed separately for each chunk at 30 frames per second for all compared methods.
+We use the dataset labels and model estimations to calculate several evaluation metrics for each chunk:
+Mean absolute error (MAE, smaller is better), signal-to-noise ratio (SNR, larger is better), and pearson correlation coefficient ($r$, larger is better).
+We then report the mean for each of these metrics across all chunks.
 
-\subsection{Which factors impact estimation performance?}
+\newcommand{\printMetric}[1]{%
+  \ifthenelse{\equal{#1}{0.0}}%
+    {--}%
+    {\sisetup{round-mode=places,round-precision=2}\num{#1}}%
+}
 
-Regression analysis to determine factors impacting estimation performance
+\begin{table}[h!]
+ 	\caption{Vitals estimation results on VV-Medium}
+ 	\label{tab:results-vv-medium}
+ 	\centering
+ 	\begin{threeparttable}	
+  	\csvreader[
+		tabular=lcccccc,
+    	table head=\toprule Method & \multicolumn{1}{c}{Heart rate} & \multicolumn{2}{c}{Pulse wave} & \multicolumn{1}{c}{Respiratory rate\tnote{a}} & \multicolumn{2}{c}{Respiration wave\tnote{a}} \\ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ \\ \midrule,
+    	table foot=\bottomrule,
+    	late after line=\\,
+    	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
+  	]{data/results_vv_medium.csv}{}%
+  {\csvcoli & \printMetric{\csvcolii} & \printMetric{\csvcoliii} & \printMetric{\csvcoliv} & \printMetric{\csvcolv} & \printMetric{\csvcolvi} & \printMetric{\csvcolvii} }
+  \begin{tablenotes}
+	\item[a] Missing respiratory signals were synthetically created and manually verified.
+	\end{tablenotes}
+  \end{threeparttable}
+\end{table}
+
+\begin{table}[h!]
+ 	\caption{Vitals estimation results on PROSIT test set}
+ 	\label{tab:results-prosit-test}
+ 	\centering
+  	\csvreader[
+		tabular=lcccccc,
+    	table head=\toprule Method & \multicolumn{1}{c}{Heart rate} & \multicolumn{2}{c}{Pulse wave} & \multicolumn{1}{c}{Respiratory rate} & \multicolumn{2}{c}{Respiration wave} \\ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ \\ \midrule,
+    	table foot=\bottomrule,
+    	late after line=\\,
+    	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
+  	]{data/results_prosit_test.csv}{}%
+  {\csvcoli & \printMetric{\csvcolii} & \printMetric{\csvcoliii} & \printMetric{\csvcoliv} & \printMetric{\csvcolv} & \printMetric{\csvcolvi} & \printMetric{\csvcolvii} }
+\end{table}
+
+\begin{table}[h!]
+ 	\caption{Vitals estimation results on VV-Ghana-Small test set}
+ 	\label{tab:results-vv-ghana-small-test}
+ 	\centering
+ 	\begin{threeparttable}
+  	\csvreader[
+		tabular=lcccccc,
+    	table head=\toprule Method & \multicolumn{1}{c}{Heart rate} & \multicolumn{2}{c}{Pulse wave} & \multicolumn{1}{c}{Respiratory rate\tnote{a}} & \multicolumn{2}{c}{Respiration wave\tnote{a}} \\ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ \\ \midrule,
+    	table foot=\bottomrule,
+    	late after line=\\,
+    	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
+  	]{data/results_vv_ghana_test.csv}{}%
+  {\csvcoli & \printMetric{\csvcolii} & \printMetric{\csvcoliii} & \printMetric{\csvcoliv} & \printMetric{\csvcolv} & \printMetric{\csvcolvi} & \printMetric{\csvcolvii} }
+  \begin{tablenotes}
+	\item[a] Missing respiratory signals were synthetically created and manually verified.
+	\end{tablenotes}
+  \end{threeparttable}
+\end{table}
+
+\subsection{Which factors impact heart rate estimation performance?}
+
+We consider a number of factors that may impact heart rate estimation performance:
+
+\begin{itemize}
+	\item \texttt{age}: The age of the participant
+	\item \texttt{camera\_stationary}: Indicates whether the camera is stationary.
+	\item \texttt{gender\_male}: Dummy variable indicating whether the participant is male.
+	\item \texttt{illuminance\_var}: Measures how much the illuminance of the participant's faces varies throughout a chunk, in interval [0,1].
+	\item \texttt{movement}: Measures how much the participant moved throughout the chunk, in interval [0,1].
+	\item \texttt{skin\_type\_x}: Dummy variable indicating whether the participant has skin type x. Base case is skin type 1.
+\end{itemize}
+
+To investigate which of these factors have a significant impact, we conduct OLS regressions using them to predict the heart rate estimation MAE.
+We perform separate regressions using the chunks of the VV-Medium dataset and the PROSIT test set - the results are reported in Tables \ref{tab:vv-hr-regression-vv} and \ref{tab:vv-hr-regression-prosit}.
+
+\begin{table}[h!]
+\begin{center}
+\caption{Regression analysis of factors affecting heart rate estimation on VV-Medium}
+\label{tab:vv-hr-regression-vv}
+\begin{tabular}{lclc}
+\toprule
+\textbf{Dep. Variable:}          &  HR MAE          & \textbf{  R-squared:         } &     0.075   \\
+\textbf{Model:}                  &       OLS        & \textbf{  Adj. R-squared:    } &     0.068   \\
+\textbf{Method:}                 &  Least Squares   & \textbf{  F-statistic:       } &     9.914   \\
+\textbf{No. Observations:}       &        1108      & \textbf{  Prob (F-statistic):} &  1.00e-14   \\
+\textbf{Df Residuals:}           &        1098      & \textbf{  AIC:               } &    -419.0   \\
+\textbf{Df Model:}               &           9      & \textbf{  BIC:               } &    -368.9   \\
+\bottomrule
+\end{tabular}
+\begin{tabular}{lcccccc}
+                                 & \textbf{coef} & \textbf{std err} & \textbf{t} & \textbf{P$> |$t$|$} & \textbf{[0.025} & \textbf{0.975]}  \\
+\midrule
+\textbf{intercept}               &       0.5320  &        0.035     &    15.257  &         0.000        &        0.464    &        0.600     \\
+\textbf{age}            			 &      -0.0014  &        0.000     &    -4.658  &         0.000        &       -0.002    &       -0.001     \\
+\textbf{illuminance\_var} 		 &       0.3539  &        0.136     &     2.594  &         0.010        &        0.086    &        0.622     \\
+\textbf{movement}       			 &       0.1967  &        0.106     &     1.855  &         0.064        &       -0.011    &        0.405     \\
+\textbf{gender\_male}   			 &       0.0098  &        0.012     &     0.803  &         0.422        &       -0.014    &        0.034     \\
+\textbf{skin\_type\_2}  			 &       0.0107  &        0.034     &     0.315  &         0.753        &       -0.056    &        0.078     \\
+\textbf{skin\_type\_3}  			 &      -0.0511  &        0.037     &    -1.390  &         0.165        &       -0.123    &        0.021     \\
+\textbf{skin\_type\_4}  			 &      -0.0565  &        0.039     &    -1.458  &         0.145        &       -0.132    &        0.020     \\
+\textbf{skin\_type\_5}  			 &       0.0164  &        0.038     &     0.433  &         0.665        &       -0.058    &        0.091     \\
+\textbf{skin\_type\_6}  			 &       0.1486  &        0.041     &     3.594  &         0.000        &        0.067    &        0.230     \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+The regression for VV-Medium shows a significant result.
+Only \texttt{age}, \texttt{illuminance\_var}, and \texttt{skin\_type\_6} are shown to have significant effects at the 5\% level:
+
+\begin{itemize}
+	\item \texttt{age} has a weak negative effect, meaning that estimations were slightly more accurate for older participants,
+	\item \texttt{illuminance\_var} had a moderate positive effect, meaning that higher variance in facial illumination of participants led to less accurate estimations, and
+	\item \texttt{skin\_type\_6} had a moderate positive effect, meaning that the error of heart rate estimation for participants with skin type 6 was on average ca.
+0.15 bpm higher, taking into account the other factors.
+\end{itemize}
+
+For VV-Medium, the factors are collectively able to explain a modest 7.5\% of the variance of the absolute errors in HR estimation.
+
+\begin{table}[h!]
+\begin{center}
+\caption{Regression analysis of factors affecting heart rate estimation on PROSIT test set}
+\label{tab:vv-hr-regression-prosit}
+\begin{tabular}{lclc}
+\toprule
+\textbf{Dep. Variable:}          &  HR MAE          & \textbf{  R-squared:         } &     0.350   \\
+\textbf{Model:}                  &       OLS        & \textbf{  Adj. R-squared:    } &     0.346   \\
+\textbf{Method:}                 &  Least Squares   & \textbf{  F-statistic:       } &     93.64   \\
+\textbf{No. Observations:}       &        1403      & \textbf{  Prob (F-statistic):} & 1.58e-124   \\
+\textbf{Df Residuals:}           &        1394      & \textbf{  AIC:               } &    -231.2   \\
+\textbf{Df Model:}               &           8      & \textbf{  BIC:               } &    -184.0   \\
+\bottomrule
+\end{tabular}
+\begin{tabular}{lcccccc}
+                                 & \textbf{coef} & \textbf{std err} & \textbf{t} & \textbf{P$> |$t$|$} & \textbf{[0.025} & \textbf{0.975]}  \\
+\midrule
+\textbf{intercept}               &       0.5039  &        0.028     &    17.823  &         0.000        &        0.448    &        0.559     \\
+\textbf{age}                     &      -0.0041  &        0.000     &    -9.169  &         0.000        &       -0.005    &       -0.003     \\
+\textbf{illuminance\_var}        &       0.6905  &        0.075     &     9.201  &         0.000        &        0.543    &        0.838     \\
+\textbf{movement}                &       0.3509  &        0.039     &     8.989  &         0.000        &        0.274    &        0.428     \\
+\textbf{camera\_stationary}      &      -0.1748  &        0.017     &   -10.056  &         0.000        &       -0.209    &       -0.141     \\
+\textbf{gender\_male}            &       0.1722  &        0.014     &    11.998  &         0.000        &        0.144    &        0.200     \\
+\textbf{skin\_type\_2}           &       0.1268  &        0.016     &     8.013  &         0.000        &        0.096    &        0.158     \\
+\textbf{skin\_type\_3}           &      -0.1226  &        0.058     &    -2.124  &         0.034        &       -0.236    &       -0.009     \\
+\textbf{skin\_type\_4}           &       0.3286  &        0.039     &     8.391  &         0.000        &        0.252    &        0.405     \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+Looking at the regression for the PROSIT test set, we find that it has much more explanatory power.
+Here, the regression itself and all considered factors appear to have significant effects at the 5\% level:
+
+\begin{itemize}
+	\item \texttt{age} has a weak negative effect, same direction but stronger than the result on VV-Medium. This means that again, estimations were slightly more accurate for older participants.
+	\item \texttt{camera\_stationary} had a negative effect, meaning that estimations were better when the camera was stationary.
+	\item \texttt{gender\_male} has a moderately positive effect, meaning that estimations were worse for male participants.
+	\item \texttt{illuminance\_var} had a strong positive effect, meaning that estimations were worse with higher variance in facial illumination of participants.
+	\item \texttt{movement} had a positive effect, meaning that estimations were worse when participants moved more.
+	\item \texttt{skin\_type\_x}: Only skin types 1, 2, 3, and 4 were present in the PROSIT test set and the result is mixed, so there is no clear conclusion.
+\end{itemize}
+
+For PROSIT, the factors are collectively able to explain a respectable 35\% of the variance of the absolute errors in HR estimation.
+
+These results confirm that both participant movement and variation in the illumination of the participant's face have a negative impact on estimation performance.
+However, we note that these effects are contained to less than 1 bpm from the best to worst values of the respective factors.
+Secondly, it is interesting to note that at least for these datasets and the our estimation model, the effect of illumination variation seems to have a greater impact than participant movement.
+Additionally, we observe that estimation performance appears to be better for female participants.
+On the one hand, this may be influenced by bearded male participants, which we don't control for -- however, on the other hand, we also don't control for use of make-up by female participants.
+
+Based on these results, we advise users of VitalLens that for best performance they should (i) hold still and place the device on a surface, and (ii) make sure the face is evenly illuminated.
+
+\subsection{Which factors impact respiratory rate estimation performance?}
 
@@
 
+\subsection{Impact of vitals}
+
+Investigate whether predictions are worse for lower / higher HR
+
 \section{Conclusion}

### 2023-12-08 paper.py

@@ -1,18 +1,20 @@
 import numpy as np
 import pandas as pd
+import statsmodels.api as sm
 
 # == Import data from csv ==
 
 # PROSIT
-chunks_prosit = pd.read_csv("export_prosit.csv")[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'env_illuminance', 'env_illuminance_d', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate']]
+chunks_prosit = pd.read_csv("raw_data/export_prosit.csv")[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'env_illuminance', 'env_illuminance_d', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate']]
 chunks_prosit['subject_id'] = chunks_prosit['session_description'].str.split('-').str.get(0)
 sessions_prosit = chunks_prosit.drop_duplicates(subset='session_id')
+subjects_prosit = sessions_prosit.drop_duplicates(subset='subject_id')
 # VV
[…]

### 2023-12-08 paper.tex
Regression analysis for RR

@@
 As part of this step, we create small chunks, calculate summary vitals, and extract further metadata.
-Note that for many chunks, the missing respiratory signal was synthetically created with an older version of VitalLens and then manually verified.
+Note that for most chunks, the missing respiratory signal was synthetically created.\footnote{This was done using an earlier version of our model trained on PROSIT.
+We then manually verified the correctness by visual inspection of both the label and video, and discarded bad labels.}
 
@@
 As part of this step, we create small chunks, calculate summary vitals, and extract further metadata.
-Note that for some chunks, the missing respiratory signal was synthetically created with an older version of VitalLens and then manually verified.
+Note that for some chunks, the missing respiratory signal was synthetically created using the same procedure as for VV-Medium.
 
@@
 
+For our analysis of the impact on respiratory rate estimation, we consider the factors \texttt{age}, \texttt{camera\_stationary}, \texttt{gender\_male}, \texttt{illuminance\_var}, and \texttt{movement}.
+We conduct another OLS regressions on the PROSIT test set using these factors to predict the respiratory rate estimation MAE - the results are reported in Table \ref{tab:vv-rr-regression-prosit}.
+
+\begin{table}[h!]
+\begin{center}
+\caption{Regression analysis of factors affecting respiratory rate estimation on PROSIT test set}
+\label{tab:vv-rr-regression-prosit}
+\begin{tabular}{lclc}
+\toprule
+\textbf{Dep. Variable:}          &  resp\_mae\_vl   & \textbf{  R-squared:         } &     0.126   \\
+\textbf{Model:}                  &       OLS        & \textbf{  Adj. R-squared:    } &     0.122   \\
+\textbf{Method:}                 &  Least Squares   & \textbf{  F-statistic:       } &     40.11   \\
+\textbf{No. Observations:}       &        1403      & \textbf{  Prob (F-statistic):} &  1.28e-38   \\
+\textbf{Df Residuals:}           &        1397      & \textbf{  AIC:               } &     396.3   \\
+\textbf{Df Model:}               &           5      & \textbf{  BIC:               } &     427.8   \\
+\bottomrule
+\end{tabular}
+\begin{tabular}{lcccccc}
+                                 & \textbf{coef} & \textbf{std err} & \textbf{t} & \textbf{P$> |$t$|$} & \textbf{[0.025} & \textbf{0.975]}  \\
+\midrule
+\textbf{intercept}               &       0.7648  &        0.031     &    24.842  &         0.000        &        0.704    &        0.825     \\
+\textbf{age}                     &      -0.0025  &        0.001     &    -4.849  &         0.000        &       -0.004    &       -0.001     \\
+\textbf{illuminance\_var}        &       0.5166  &        0.091     &     5.682  &         0.000        &        0.338    &        0.695     \\
+\textbf{movement}                &       0.5376  &        0.048     &    11.210  &         0.000        &        0.444    &        0.632     \\
+\textbf{camera\_stationary}      &      -0.0351  &        0.020     &    -1.731  &         0.084        &       -0.075    &        0.005     \\
+\textbf{gender\_male}            &      -0.0076  &        0.016     &    -0.481  &         0.630        &       -0.039    &        0.023     \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+The regression itself and all the factors \texttt{age}, \texttt{illuminance\_var}, and \texttt{movement} have significant effects at the 5\% level.
+Contrary to expectations, \texttt{camera\_stationary} does not have a significant effect, only significant at 8.5\%.
+
+\begin{itemize}
+	\item \texttt{age} has a weak negative effect, indicating that estimations were slightly more accurate for older participants.
+	\item \texttt{illuminance\_var} had a positive effect, meaning that estimations were worse with higher variance in facial illumination of participants.
+	\item \texttt{movement} had a positive effect, meaning that estimations were worse when participants moved more.
+\end{itemize}
+
+The factors are collectively able to explain 12.6\% of the variance of the absolute errors in RR estimation.
+As expected, participant movement degrades the estimation quality for the respiratory rate.
+It is surprising that variation in the illumination of the participant's face impacts the estimation to a similar degree, and that this effect is stronger and more significant than the stationarity of the camera.
+
+These results reinforce the advice given to users of VitalLens.
+
 \subsection{Impact of subject movement}

### 2023-12-08 paper.py

@@ -207,15 +207,7 @@ results_prosit_test_summary.to_csv("data/results_prosit_test.csv", index=False)
 categorical_cols = ['camera_stationary', 'subject_gender', 'subject_skin_type']
 numerical_cols = ['subject_age', 'subject_illuminance_d', 'subject_movement']
 
-## VV-Medium - SNR
-# vv_medium_pulse_snr_reg = results_vv_medium[categorical_cols + numerical_cols + ['pulse_snr_vl']].dropna()
-# X_categorical = pd.get_dummies(vv_medium_pulse_snr_reg[categorical_cols], columns=categorical_cols, drop_first=True)
-# X = pd.concat([vv_medium_pulse_snr_reg[numerical_cols], X_categorical], axis=1).astype('float64')
-# y = vv_medium_pulse_snr_reg['pulse_snr_vl'].astype(float)
-# model = sm.OLS(y, X).fit()
-# print("VV-Medium")
-# print(model.summary())
-## VV-Medium - MAE
+## VV-Medium - HR MAE
 vv_medium_pulse_mae_reg = results_vv_medium[categorical_cols + numerical_cols + ['pulse_mae_vl']].dropna()
 X_categorical = pd.get_dummies(vv_medium_pulse_mae_reg[categorical_cols], columns=categorical_cols, drop_first=True)
 X = pd.concat([vv_medium_pulse_mae_reg[numerical_cols], X_categorical], axis=1).astype('float64')
@@ -226,20 +218,26 @@ print("VV-Medium")
 print(model.summary())
 print(model.summary().as_latex())
 
-## PROSIT - SNR
-# prosit_pulse_snr_reg = results_prosit_test[categorical_cols + numerical_cols + ['pulse_snr_vl']].dropna()
-# X_categorical = pd.get_dummies(prosit_pulse_snr_reg[categorical_cols], columns=categorical_cols, drop_first=True)
-# X = pd.concat([prosit_pulse_snr_reg[numerical_cols], X_categorical], axis=1).astype('float64')
-# y = prosit_pulse_snr_reg['pulse_snr_vl'].astype(float)
-# model = sm.OLS(y, X).fit()
-# print("VV-Medium")
-# print(model.summary())
-## PROSIT - MAE
+## PROSIT - HR MAE
 prosit_pulse_mae_reg = results_prosit_test[categorical_cols + numerical_cols + ['pulse_mae_vl']].dropna()
[…]

### 2023-12-11 paper.tex
Finish first draft

--- 2023-12-08 paper.tex
@@
 \usepackage[flushleft]{threeparttable} % For tablenotes
+\usepackage{pgfplotstable}
+\usepackage{enumitem}
 
@@
 
+\csvreader[before reading=\def\totalparticipantsprosit{0}\def\totalchunksprosit{0}\def\totaltimeprosit{0}\def\totalsessionsprosit{0}]{data/prosit_summary.csv}{participants=\participants,chunks=\chunks,time=\time,sessions=\sessions}{%
+	\pgfmathsetmacro{\totalparticipantsprosit}{\totalparticipantsprosit+\participants}%
+	\pgfmathsetmacro{\totalchunksprosit}{\totalchunksprosit+\chunks}%
+	\pgfmathsetmacro{\totaltimeprosit}{\totaltimeprosit+\time}%
+	\pgfmathsetmacro{\totalsessionsprosit}{\totalsessionsprosit+\sessions}%
+}
+
+\csvreader[before reading=\def\totalparticipantsvvmedium{0}\def\totalchunksvvmedium{0}\def\totaltimevvmedium{0}]{data/vv_medium_summary.csv}{participants=\participants,chunks=\chunks,time=\time}{%
+	\pgfmathsetmacro{\totalparticipantsvvmedium}{\totalparticipantsvvmedium+\participants}%
+	\pgfmathsetmacro{\totalchunksvvmedium}{\totalchunksvvmedium+\chunks}%
+	\pgfmathsetmacro{\totaltimevvmedium}{\totaltimevvmedium+\time}%
+}
+
+\csvreader[before reading=\def\totalparticipantsvvghana{0}\def\totalchunksvvghana{0}\def\totaltimevvghana{0}]{data/vv_ghana_small_summary.csv}{participants=\participants,chunks=\chunks,time=\time}{%
+	\pgfmathsetmacro{\totalparticipantsvvghana}{\totalparticipantsvvghana+\participants}%
+	\pgfmathsetmacro{\totalchunksvvghana}{\totalchunksvvghana+\chunks}%
+	\pgfmathsetmacro{\totaltimevvghana}{\totaltimevvghana+\time}%
+}
+
 \begin{document}
@@
 This report introduces VitalLens, an app that estimates vital signs such as heart rate and respiration rate from selfie video in real time.
-The estimation engine of VitalLens is a computer vision model trained on a diverse dataset of video and gold-standard ground truth physiological sensor data.
-% TODO: Summarize findings
+VitalLens uses a computer vision model trained on a diverse dataset of video and physiological sensor data.
+We benchmark performance on several diverse datasets, including VV-Medium, which consists of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvmedium} unique participants.
+VitalLens outperforms several existing methods including POS and MTTS-CAN on all datasets while maintaining a fast inference speed.
+On VV-Medium, VitalLens achieves absolute errors of 0.71 bpm for heart rate estimation, and 0.76 rpm for respiratory rate estimation.
 \end{abstract}
@@
 Given enough high-quality data to learn from, learning-based approaches such as DeepPhys \cite{chen2018deep} and MTTS-CAN \cite{liu2020multi} hold the promise of greater accuracy, which they have to trade off against inference speed.
-There are also other learning-based approaches achieving high accuracy, which we do not cover here as they are not aiming to achieve real-time inference.
+There are also other learning-based approaches achieving higher accuracy, which we do not cover here as they are not aiming to achieve real-time inference.
 
 In this context, we introduce VitalLens, an innovative application designed to estimate vital signs, including heart rate and respiratory rate, in real time from selfie videos.
-The underlying estimation engine of VitalLens is built upon a computer vision model trained on a diverse dataset, combining video recordings with gold-standard ground truth physiological sensor data.
+This report focuses on the capabilities and limitation of VitalLens.
+Our key contributions are:
 
-[INSERT FIGURE WITH SAMPLE PREDICTIONS FROM VALIDATION SET]
-
-This report focuses on the capabilities and limitation of VitalLens.
-To this end, We present a comprehensive performance evaluation, comparing it with the named existing methods on diverse datasets.
-
-The remainder of this report is organized as follows: Section \ref{sec:scope} provides an brief overview of the model, Section \ref{sec:datasets} details the datasets used for training and evaluation, Section \ref{sec:methodology} outlines the systematic approach to model development and evaluation, Section \ref{sec:results} presents the results and discussions, and finally, Section \ref{sec:conclusion} concludes the report with implications and future directions.
+\begin{itemize}
+	\item \textbf{Real-time rPPG application.} We introduce VitalLens -- the first widely distributed rPPG application for real time estimation of vitals such as of heart rate and respiratory rate. 
+	\item \textbf{Comprehensive evaluation on a large dataset.} We evaluate the performance of VitalLens on multiple datasets, including the large VV-Medium \cite{toye2023vital} with \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvmedium} participants.
+VitalLens outperforms G, CHROM, POS, DeepPhys, and MTTS-CAN in terms of estimation accuracy.
+On VV-Medium, VitalLens achieves absolute errors of 0.71 bpm for heart rate estimation, and 0.76 rpm for respiratory rate estimation, while taking 18 ms to run inference.
+	\item \textbf{Analysis of factors impacting performance.} We systematically investigate factors impacting estimation performance.
+We find that \textit{participant movement} and \textit{variation of participant illuminance} are the main factors impacting performance.
+Compared to approaches like POS, VitalLens is less impacted by darker skin types when estimating heart rates.
+	\item \textbf{Maintaining end-user privacy}. VitalLens runs inference locally, meaning that video and vitals never leave the device.
+\end{itemize}
 
@@
 
-The estimation engine built into VitalLens is a computer vision model trained on a diverse dataset of video and gold-standard ground truth physiological data.
+The estimation engine built into VitalLens (henceforth ``VitalLens'') is a computer vision model trained on a diverse dataset of video and gold-standard ground truth physiological data.
 The model architecture is broadly based on the \textit{EfficientNetV2} \cite{tan2021efficient} model family, enhanced with several improvements in architecture and model optimization to enable efficient training and inference in the rPPG domain.
@@
 We pre-process and split each session into small chunks of 5-20 seconds with valid video and signals.
-As part of this step, we also calculate summary vitals for each chunk from the continuous signals, and extract further metadata such as the amount of participant movement and illumination variation.
+As part of this step, we also calculate summary vitals for each chunk from the continuous signals, and extract further metadata such as the amount of participant movement and illuminance variation.
 
 \paragraph{Dataset size and split.}
-\csvreader[before reading=\def\totalparticipantsprosit{0}\def\totalchunksprosit{0}\def\totaltimeprosit{0}\def\totalsessionsprosit{0}]{data/prosit_summary.csv}{participants=\participants,chunks=\chunks,time=\time,sessions=\sessions}{%
-	\pgfmathsetmacro{\totalparticipantsprosit}{\totalparticipantsprosit+\participants}%
-	\pgfmathsetmacro{\totalchunksprosit}{\totalchunksprosit+\chunks}%
-	\pgfmathsetmacro{\totaltimeprosit}{\totaltimeprosit+\time}%
-	\pgfmathsetmacro{\totalsessionsprosit}{\totalsessionsprosit+\sessions}%
-}
 Development of PROSIT is ongoing.
@@
 It is the largest dataset we are aware of that is available for research without academic affiliation.
-\csvreader[before reading=\def\totalparticipantsvvmedium{0}\def\totalchunksvvmedium{0}\def\totaltimevvmedium{0}]{data/vv_medium_summary.csv}{participants=\participants,chunks=\chunks,time=\time}{%
-	\pgfmathsetmacro{\totalparticipantsvvmedium}{\totalparticipantsvvmedium+\participants}%
-	\pgfmathsetmacro{\totalchunksvvmedium}{\totalchunksvvmedium+\chunks}%
-	\pgfmathsetmacro{\totaltimevvmedium}{\totaltimevvmedium+\time}%
-}
 We use a slightly extended version of the medium instantiation (``VV-Medium''), which consists of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvmedium} participants.
@@
 It was collected in Ghana, with the majority of participants having skin type 5 or 6.
-\csvreader[before reading=\def\totalparticipantsvvghana{0}\def\totalchunksvvghana{0}\def\totaltimevvghana{0}]{data/vv_ghana_small_summary.csv}{participants=\participants,chunks=\chunks,time=\time}{%
-	\pgfmathsetmacro{\totalparticipantsvvghana}{\totalparticipantsvvghana+\participants}%
-	\pgfmathsetmacro{\totalchunksvvghana}{\totalchunksvvghana+\chunks}%
-	\pgfmathsetmacro{\totaltimevvghana}{\totaltimevvghana+\time}%
-}
 We use a small instantiation (``VV-Ghana-Small''), which consists of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvghana} participants.
@@
 
-To develop and evaluate the VitalLens estimation engine model, we follow a systematic approach involving model training, validation, and testing:
+To develop and evaluate VitalLens, we follow a systematic approach involving model training, validation, and testing:
 
@@
 \paragraph{Model Testing.}
-The final evaluation of the VitalLens estimation engine is conducted by benchmarking the generalization abilities of the final model on several datasets, all of which are participant-disjoint from both the training and validation datasets:
+The final evaluation of VitalLens is conducted by benchmarking the generalization abilities of the final model on several datasets, all of which are participant-disjoint from both the training and validation datasets:
 
@@
 \end{itemize}
+
+\paragraph{Inference and Privacy.}
+When end-users use VitalLens on their devices, inference runs locally.
+This means that the video and vitals of end-users never leave their devices, which is important to maintain privacy.
+Another benefit is that an internet connection is not required to run VitalLens.
 
@@
   	\csvreader[
-		tabular=lcccccc,
-    	table head=\toprule Method & \multicolumn{1}{c}{Heart rate} & \multicolumn{2}{c}{Pulse wave} & \multicolumn{1}{c}{Respiratory rate\tnote{a}} & \multicolumn{2}{c}{Respiration wave\tnote{a}} \\ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ \\ \midrule,
+		tabular=lccccccc,
+    	table head=\toprule Method & \multicolumn{1}{c}{Heart rate} & \multicolumn{2}{c}{Pulse wave} & \multicolumn{1}{c}{Respiratory rate\tnote{a}} & \multicolumn{2}{c}{Respiration wave\tnote{a}} & Inference \\ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & speed in ms\tnote{b} \\ \midrule,
     	table foot=\bottomrule,
@@
   	]{data/results_vv_medium.csv}{}%
-  {\csvcoli & \printMetric{\csvcolii} & \printMetric{\csvcoliii} & \printMetric{\csvcoliv} & \printMetric{\csvcolv} & \printMetric{\csvcolvi} & \printMetric{\csvcolvii} }
+  {\csvcoli & \printMetric{\csvcolii} & \printMetric{\csvcoliii} & \printMetric{\csvcoliv} & \printMetric{\csvcolv} & \printMetric{\csvcolvi} & \printMetric{\csvcolvii} & \sisetup{round-mode=places,round-precision=1}\num{\csvcolviii}}
   \begin{tablenotes}
 	\item[a] Missing respiratory signals were synthetically created and manually verified.
+	\item[b] Time to run inference for a single video frame.
 	\end{tablenotes}
@@
 \end{table}
+
+Results for VV-Medium are reported in Table \ref{tab:results-vv-medium}.
+The heart rate and respiratory rate estimated by VitalLens across all chunks was off by 0.71 bpm and 0.76 rpm on average from the gold-standard labels, while taking 18 ms for inference.
+VitalLens beats all other listed methods for all considered evaluation metrics, while maintaining a competitive inference speed that is sufficient to run in real time at 30 fps.\footnote{The inference speeds are for a single video frame, excluding face detection but including all other computation, such as resizing.
+We measure the inference speeds using our own implementations as an average across 1,000 iterations on a workstation with Nvidia GeForce RTX 3080 and AMD Ryzen 7 3700X.}
+The next best predictions for heart rate and respiratory rate also came from learning-based models MTTS-CAN and DeepPhys, respectively.
+Note that this dataset has a large variety of participants and their demographics, but low amounts of participant and camera movement.
 
@@
 \end{table}
+
+The PROSIT test set has a lower variety in participants and their demographics, but it includes both participant and camera movement.
+Results for the PROSIT test set are reported in Table \ref{tab:results-prosit-test} -- again, VitalLens beats all listed methods for all considered evaluation metrics.
+However, we observe that here, the absolute errors achieved by VitalLens for heart rate and respiratory rate are 4.62 bpm and 3.76, respectively -- markedly higher than for VV-Medium.
+This is a first hint that factors like participant and camera movement are the main drivers of lower estimation accuracies.
 
@@
 
-\subsection{Which factors impact heart rate estimation performance?}
+Finally, Table \ref{tab:results-vv-ghana-small-test} lists the results for the VV-Ghana-Small test set.
+The results paint a similar picture as the two previous datasets, with VitalLens beating out the other methods.
+In comparison to VV-Medium, we observe that the errors are generally higher - in this case, this is caused by the demographic differences between the participants.
+We further investigate this in the following sections.
 
-We consider a number of factors that may impact heart rate estimation performance:
+\subsection{Which factors impact estimation performance?}
 
-\begin{itemize}
-	\item \texttt{age}: The age of the participant
-	\item \texttt{camera\_stationary}: Indicates whether the camera is stationary.
-	\item \texttt{gender\_male}: Dummy variable indicating whether the participant is male.
-	\item \texttt{illuminance\_var}: Measures how much the illuminance of the participant's faces varies throughout a chunk, in interval [0,1].
-	\item \texttt{movement}: Measures how much the participant moved throughout the chunk, in interval [0,1].
-	\item \texttt{skin\_type\_x}: Dummy variable indicating whether the participant has skin type x. Base case is skin type 1.
-\end{itemize}
+We consider a number of factors that may impact estimation performance of pulse and respiration.
 
-To investigate which of these factors have a significant impact, we conduct OLS regressions using them to predict the heart rate estimation MAE.
-We perform separate regressions using the chunks of the VV-Medium dataset and the PROSIT test set - the results are reported in Tables \ref{tab:vv-hr-regression-vv} and \ref{tab:vv-hr-regression-prosit}.
+\paragraph{Demographic factors:}
+
+\begin{enumerate}[align=left]
+	\item[\texttt{age}:] The age of the participant.
+	\item[\texttt{gender\_male}:] Dummy variable indicating whether the participant is male.
+	\item[\texttt{hr} / \texttt{rr}:] The heart rate and respiratory rate of the participant.
+	\item[\texttt{skin\_type\_x}:] Dummy variable indicating whether the participant has skin type x. Base case is skin type 1.
+\end{enumerate}
+
+\paragraph{Behavioral/environmental factors:}
+
+\begin{enumerate}[align=left]
+	\item[\texttt{camera\_stationary}:] Indicates whether the camera is stationary.
+	\item[\texttt{illuminance\_var}:] Measures how much the illuminance of the participant's faces varies throughout a chunk, in interval [0,1].
+Derived from the luma component of the pixels within the participant's face.
+Partially caused by movement ($r = 0.2$ for PROSIT).
+	\item[\texttt{movement}:] Measures how much the participant moved throughout the chunk, in interval [0,1]. Derived from the movement of facial landmarks.
+\end{enumerate}
+
+\subsubsection{Pulse wave}
+
+To investigate which factors have a significant impact, we conduct OLS regressions using them to predict the SNR for pulse wave estimation.
+We perform two separate regressions, one using the chunks of the VV-Medium dataset and one for the PROSIT test set - the results are reported in Tables \ref{tab:vv-hr-regression-vv} and \ref{tab:vv-hr-regression-prosit}.
 
@@
 \begin{center}
-\caption{Regression analysis of factors affecting heart rate estimation on VV-Medium}
+\caption{Regression analysis of factors affecting pulse wave estimation on VV-Medium}
 \label{tab:vv-hr-regression-vv}
@@
 \toprule
-\textbf{Dep. Variable:}          &  HR MAE          & \textbf{  R-squared:         } &     0.075   \\
-\textbf{Model:}                  &       OLS        & \textbf{  Adj. R-squared:    } &     0.068   \\
-\textbf{Method:}                 &  Least Squares   & \textbf{  F-statistic:       } &     9.914   \\
-\textbf{No. Observations:}       &        1108      & \textbf{  Prob (F-statistic):} &  1.00e-14   \\
-\textbf{Df Residuals:}           &        1098      & \textbf{  AIC:               } &    -419.0   \\
-\textbf{Df Model:}               &           9      & \textbf{  BIC:               } &    -368.9   \\
+\textbf{Dep. Variable:}          &  Pulse wave SNR  & \textbf{  R-squared:         } &     0.264   \\
+\textbf{Model:}                  &       OLS        & \textbf{  Adj. R-squared:    } &     0.257   \\
+\textbf{Method:}                 &  Least Squares   & \textbf{  F-statistic:       } &     39.27   \\
+\textbf{No. Observations:}       &        1108      & \textbf{  Prob (F-statistic):} &  2.40e-66   \\
+\textbf{Df Residuals:}           &        1097      & \textbf{  AIC:               } &     7118.   \\
+\textbf{Df Model:}               &          10      & \textbf{  BIC:               } &     7174.   \\
 \bottomrule
@@
 \midrule
-\textbf{intercept}               &       0.5320  &        0.035     &    15.257  &         0.000        &        0.464    &        0.600     \\
-\textbf{age}            			 &      -0.0014  &        0.000     &    -4.658  &         0.000        &       -0.002    &       -0.001     \\
-\textbf{illuminance\_var} 		 &       0.3539  &        0.136     &     2.594  &         0.010        &        0.086    &        0.622     \\
-\textbf{movement}       			 &       0.1967  &        0.106     &     1.855  &         0.064        &       -0.011    &        0.405     \\
-\textbf{gender\_male}   			 &       0.0098  &        0.012     &     0.803  &         0.422        &       -0.014    &        0.034     \\
-\textbf{skin\_type\_2}  			 &       0.0107  &        0.034     &     0.315  &         0.753        &       -0.056    &        0.078     \\
-\textbf{skin\_type\_3}  			 &      -0.0511  &        0.037     &    -1.390  &         0.165        &       -0.123    &        0.021     \\
-\textbf{skin\_type\_4}  			 &      -0.0565  &        0.039     &    -1.458  &         0.145        &       -0.132    &        0.020     \\
-\textbf{skin\_type\_5}  			 &       0.0164  &        0.038     &     0.433  &         0.665        &       -0.058    &        0.091     \\
-\textbf{skin\_type\_6}  			 &       0.1486  &        0.041     &     3.594  &         0.000        &        0.067    &        0.230     \\
+\textbf{intercept}               &       9.7449  &        1.575     &     6.189  &         0.000        &        6.655    &       12.834     \\
+\textbf{age}                     &       0.1135  &        0.010     &    11.919  &         0.000        &        0.095    &        0.132     \\
+\textbf{illuminance\_var}        &     -22.3650  &        4.096     &    -5.460  &         0.000        &      -30.403    &      -14.327     \\
+\textbf{movement}                &      -5.3373  &        3.179     &    -1.679  &         0.093        &      -11.576    &        0.901     \\
+\textbf{hr}                      &      -0.0283  &        0.014     &    -2.046  &         0.041        &       -0.055    &       -0.001     \\
+\textbf{gender\_male}            &       0.2530  &        0.377     &     0.672  &         0.502        &       -0.486    &        0.992     \\
+\textbf{skin\_type\_2}  			 &      -0.1885  &        1.024     &    -0.184  &         0.854        &       -2.199    &        1.822     \\
+\textbf{skin\_type\_3}  			 &       0.4203  &        1.104     &     0.381  &         0.703        &       -1.745    &        2.586     \\
+\textbf{skin\_type\_4}  			 &       1.4931  &        1.161     &     1.286  &         0.199        &       -0.785    &        3.772     \\
+\textbf{skin\_type\_5}  			 &      -2.6020  &        1.140     &    -2.282  &         0.023        &       -4.839    &       -0.365     \\
+\textbf{skin\_type\_6}  			 &      -7.9805  &        1.241     &    -6.429  &         0.000        &      -10.416    &       -5.545     \\
 \bottomrule
@@
 
-The regression for VV-Medium shows a significant result.
-Only \texttt{age}, \texttt{illuminance\_var}, and \texttt{skin\_type\_6} are shown to have significant effects at the 5\% level:
+For our regression analysis on the VV-Medium, we use all available factors.
+The regression equation is significant and explains 26.4\% of the variance of the pulse wave estimation SNR.
+\texttt{age}, \texttt{hr}, \texttt{illuminance\_var}, and \texttt{skin\_type\_x} are shown to have significant effects at the 5\% level:
 
 \begin{itemize}
-	\item \texttt{age} has a weak negative effect, meaning that estimations were slightly more accurate for older participants,
-	\item \texttt{illuminance\_var} had a moderate positive effect, meaning that higher variance in facial illumination of participants led to less accurate estimations, and
-	\item \texttt{skin\_type\_6} had a moderate positive effect, meaning that the error of heart rate estimation for participants with skin type 6 was on average ca.
-0.15 bpm higher, taking into account the other factors.
+	\item \texttt{age} has a weak positive effect, meaning that estimations were slightly more accurate for older participants,
+	\item \texttt{hr} has a weak negative effect, meaning that estimations were slightly less accurate for participants with higher heart rates,
+	\item \texttt{illuminance\_var} had a strong negative effect, meaning that higher variance in facial illuminance of participants led to less accurate estimations, and
+	\item \texttt{skin\_type\_5} and especially \texttt{skin\_type\_6} had negative effects, meaning that pulse wave estimation for participants with skin types 5 and especially 6 was less accurate.
 \end{itemize}
 
-For VV-Medium, the factors are collectively able to explain a modest 7.5\% of the variance of the absolute errors in HR estimation.
+It is worth noting that \texttt{movement} did not have a significant effect, potentially because participants generally do not move much for this dataset.
 
@@
 \begin{center}
-\caption{Regression analysis of factors affecting heart rate estimation on PROSIT test set}
+\caption{Regression analysis of factors affecting pulse wave estimation on PROSIT test set}
 \label{tab:vv-hr-regression-prosit}
@@
 \toprule
-\textbf{Dep. Variable:}          &  HR MAE          & \textbf{  R-squared:         } &     0.350   \\
-\textbf{Model:}                  &       OLS        & \textbf{  Adj. R-squared:    } &     0.346   \\
-\textbf{Method:}                 &  Least Squares   & \textbf{  F-statistic:       } &     93.64   \\
-\textbf{No. Observations:}       &        1403      & \textbf{  Prob (F-statistic):} & 1.58e-124   \\
-\textbf{Df Residuals:}           &        1394      & \textbf{  AIC:               } &    -231.2   \\
-\textbf{Df Model:}               &           8      & \textbf{  BIC:               } &    -184.0   \\
+\textbf{Dep. Variable:}          &  Pulse wave SNR  & \textbf{  R-squared:         } &     0.103   \\
+\textbf{Model:}                  &       OLS        & \textbf{  Adj. R-squared:    } &     0.102   \\
+\textbf{Method:}                 &  Least Squares   & \textbf{  F-statistic:       } &     53.82   \\
+\textbf{No. Observations:}       &        1403      & \textbf{  Prob (F-statistic):} &  6.40e-33   \\
+\textbf{Df Residuals:}           &        1399      & \textbf{  AIC:               } & 1.040e+04   \\
+\textbf{Df Model:}               &           3      & \textbf{  BIC:               } & 1.042e+04   \\
 \bottomrule
@@
 \midrule
-\textbf{intercept}               &       0.5039  &        0.028     &    17.823  &         0.000        &        0.448    &        0.559     \\
-\textbf{age}                     &      -0.0041  &        0.000     &    -9.169  &         0.000        &       -0.005    &       -0.003     \\
-\textbf{illuminance\_var}        &       0.6905  &        0.075     &     9.201  &         0.000        &        0.543    &        0.838     \\
-\textbf{movement}                &       0.3509  &        0.039     &     8.989  &         0.000        &        0.274    &        0.428     \\
-\textbf{camera\_stationary}      &      -0.1748  &        0.017     &   -10.056  &         0.000        &       -0.209    &       -0.141     \\
-\textbf{gender\_male}            &       0.1722  &        0.014     &    11.998  &         0.000        &        0.144    &        0.200     \\
-\textbf{skin\_type\_2}           &       0.1268  &        0.016     &     8.013  &         0.000        &        0.096    &        0.158     \\
-\textbf{skin\_type\_3}           &      -0.1226  &        0.058     &    -2.124  &         0.034        &       -0.236    &       -0.009     \\
-\textbf{skin\_type\_4}           &       0.3286  &        0.039     &     8.391  &         0.000        &        0.252    &        0.405     \\
+\textbf{intercept}               &       7.2056  &        0.669     &    10.775  &         0.000        &        5.894    &        8.517     \\
+\textbf{illuminance\_var} 		 &     -26.4655  &        3.216     &    -8.230  &         0.000        &      -32.774    &      -20.157     \\
+\textbf{movement}       			 &     -14.3836  &        1.687     &    -8.525  &         0.000        &      -17.693    &      -11.074     \\
+\textbf{camera\_stationary}   	 &       0.1855  &        0.597     &     0.311  &         0.756        &       -0.985    &        1.356     \\
 \bottomrule
@@
 
-Looking at the regression for the PROSIT test set, we find that it has much more explanatory power.
-Here, the regression itself and all considered factors appear to have significant effects at the 5\% level:
+Since there are not many unique participants in the PROSIT test set, we only consider the behavioral and environmental factors for the next regression.
+Looking at the results, the regression equation itself as well as \texttt{illuminance\_var} and \texttt{movement} have significant effects at the 5\% level:
 
 \begin{itemize}
-	\item \texttt{age} has a weak negative effect, same direction but stronger than the result on VV-Medium. This means that again, estimations were slightly more accurate for older participants.
-	\item \texttt{camera\_stationary} had a negative effect, meaning that estimations were better when the camera was stationary.
-	\item \texttt{gender\_male} has a moderately positive effect, meaning that estimations were worse for male participants.
-	\item \texttt{illuminance\_var} had a strong positive effect, meaning that estimations were worse with higher variance in facial illumination of participants.
-	\item \texttt{movement} had a positive effect, meaning that estimations were worse when participants moved more.
-	\item \texttt{skin\_type\_x}: Only skin types 1, 2, 3, and 4 were present in the PROSIT test set and the result is mixed, so there is no clear conclusion.
+	\item \texttt{illuminance\_var} has a strong negative effect, meaning that estimations became less accurate with higher variance in facial illuminance of participants.
+	\item \texttt{movement} has a negative effect, meaning that estimations were less accurate when participants moved more.
 \end{itemize}
 
-For PROSIT, the factors are collectively able to explain a respectable 35\% of the variance of the absolute errors in HR estimation.
+These behavioral factors are collectively able to explain 10.3\% of the variance of the pulse wave estimation SNR.
+It is worth noting that \texttt{camera\_stationary} did not have a significant effect, which indicates that our rPPG model has some degree of robustness against camera movement.
 
-These results confirm that both participant movement and variation in the illumination of the participant's face have a negative impact on estimation performance.
-However, we note that these effects are contained to less than 1 bpm from the best to worst values of the respective factors.
-Secondly, it is interesting to note that at least for these datasets and the our estimation model, the effect of illumination variation seems to have a greater impact than participant movement.
-Additionally, we observe that estimation performance appears to be better for female participants.
-On the one hand, this may be influenced by bearded male participants, which we don't control for -- however, on the other hand, we also don't control for use of make-up by female participants.
+Overall, the results confirm that the behavioral factors of participant movement and variation in the illuminance of the participant's face dominate the results with have a significant negative impact on estimation performance.
+It is interesting to note that at least for these datasets, the effect of illuminance variation seems to have a greater impact than participant movement.
+Our model performance appears to be slightly impacted by the heart rate itself.
+We also confirm previous findings that rPPG models can struggle with darker skin types, but cannot confirm findings that estimation is more accurate for females \cite{nowara2020meta}.
+Additionally, we observe that estimation appears to be more accurate for older participants.
 
-Based on these results, we advise users of VitalLens that for best performance they should (i) hold still and place the device on a surface, and (ii) make sure the face is evenly illuminated.
+\subsubsection{Respiration wave}
 
-\subsection{Which factors impact respiratory rate estimation performance?}
+For our analysis of the impact on respiration wave estimation, we conduct another two OLS regressions.
+The results are reported in Table \ref{tab:vv-rr-regression-vv} and \ref{tab:vv-rr-regression-prosit}.
 
-\subsection{Impact of subject movement}
+\begin{table}[h!]
+\begin{center}
+\caption{Regression analysis of factors affecting respiratory rate estimation on VV-Medium}
+\label{tab:vv-rr-regression-vv}
+\begin{tabular}{lclc}
+\toprule
+\textbf{Dep. Variable:}          &  Resp wave SNR   & \textbf{  R-squared:         } &     0.071   \\
+\textbf{Model:}                  &       OLS        & \textbf{  Adj. R-squared:    } &     0.062   \\
+\textbf{Method:}                 &  Least Squares   & \textbf{  F-statistic:       } &     8.340   \\
+\textbf{No. Observations:}       &        1108      & \textbf{  Prob (F-statistic):} &  3.71e-13   \\
+\textbf{Df Residuals:}           &        1097      & \textbf{  AIC:               } &     7352.   \\
+\textbf{Df Model:}               &          10      & \textbf{  BIC:               } &     7407.   \\
+\bottomrule
+\end{tabular}
+\begin{tabular}{lcccccc}
+                                 & \textbf{coef} & \textbf{std err} & \textbf{t} & \textbf{P$> |$t$|$} & \textbf{[0.025} & \textbf{0.975]}  \\
+\midrule
+\textbf{intercept}               &       8.6117  &        1.349     &     6.385  &         0.000        &        5.965    &       11.258     \\
+\textbf{age}            			 &       0.0674  &        0.010     &     6.559  &         0.000        &        0.047    &        0.088     \\
+\textbf{illuminance\_var} 		 &      -4.6049  &        4.564     &    -1.009  &         0.313        &      -13.560    &        4.350     \\
+\textbf{movement}       			 &     -13.8675  &        3.536     &    -3.922  &         0.000        &      -20.805    &       -6.930     \\
+\textbf{rr}          			 &      -0.1615  &        0.040     &    -4.056  &         0.000        &       -0.240    &       -0.083     \\
+\textbf{gender\_male}   			 &       0.7883  &        0.407     &     1.939  &         0.053        &       -0.009    &        1.586     \\
+\textbf{skin\_type\_2}  			 &      -0.8223  &        1.139     &    -0.722  &         0.470        &       -3.057    &        1.413     \\
+\textbf{skin\_type\_3}  			 &      -0.5369  &        1.230     &    -0.437  &         0.663        &       -2.950    &        1.876     \\
+\textbf{skin\_type\_4}  			 &       0.1492  &        1.296     &     0.115  &         0.908        &       -2.395    &        2.693     \\
+\textbf{skin\_type\_5}  			 &       1.1431  &        1.269     &     0.901  &         0.368        &       -1.346    &        3.632     \\
+\textbf{skin\_type\_6}  			 &      -0.3998  &        1.383     &    -0.289  &         0.773        &       -3.114    &        2.314     \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 
-Bar chart comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens SNR or MAE estimating HR - vs subject movement on x axis.
-Bar chart comparing DeepPhys, MTTS-CAN, VitalLens SNR or MAE estimating RR - vs subject movement on x axis.
+For our regression analysis on VV-Medium, we use all available factors.
+The regression equation is significant but explains only 7.1\% of the variance of the respiratory wave estimation SNR.
+\texttt{age}, \texttt{movement}, and \texttt{rr} are shown to have significant effects at the 5\% level:
 
-\subsection{Impact of illumination variation}
+\begin{itemize}
+	\item \texttt{age} has a weak positive effect, meaning that estimations were slightly more accurate for older participants,
+	\item \texttt{movement} had a strong negative effect, meaning that higher participant movement led to less accurate estimations, and
+	\item \texttt{rr} has a weak negative effect, meaning that estimations were slightly less accurate for participants with higher respiratory rates.
+\end{itemize}
 
-Bar chart comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens SNR estimating HR - vs illumination variation on x axis.
+\begin{table}[h!]
+\begin{center}
+\caption{Regression analysis of factors affecting respiratory rate estimation on PROSIT test set}
+\label{tab:vv-rr-regression-prosit}
+\begin{tabular}{lclc}
+\toprule
+\textbf{Dep. Variable:}          &  Resp wave SNR   & \textbf{  R-squared:         } &     0.068   \\
+\textbf{Model:}                  &       OLS        & \textbf{  Adj. R-squared:    } &     0.066   \\
+\textbf{Method:}                 &  Least Squares   & \textbf{  F-statistic:       } &     33.96   \\
+\textbf{No. Observations:}       &        1403      & \textbf{  Prob (F-statistic):} &  3.46e-21   \\
+\textbf{Df Residuals:}           &        1399      & \textbf{  AIC:               } &     9735.   \\
+\textbf{Df Model:}               &           3      & \textbf{  BIC:               } &     9756.   \\
+\bottomrule
+\end{tabular}
+\begin{tabular}{lcccccc}
+                                 & \textbf{coef} & \textbf{std err} & \textbf{t} & \textbf{P$> |$t$|$} & \textbf{[0.025} & \textbf{0.975]}  \\
+\midrule
+\textbf{intercept}               &       2.4654  &        0.527     &     4.676  &         0.000        &        1.431    &        3.500     \\
+\textbf{illuminance\_var}  		 &     -11.0138  &        2.535     &    -4.344  &         0.000        &      -15.987    &       -6.040     \\
+\textbf{movement}                &     -11.9294  &        1.330     &    -8.968  &         0.000        &      -14.539    &       -9.320     \\
+\textbf{camera\_stationary}      &      -1.3251  &        0.471     &    -2.816  &         0.005        &       -2.248    &       -0.402     \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 
-\subsection{Impact of subject skin type}
+For the regression on the PROSIT test set, we again focus on the behavioral and environmental factors.
+The regression itself and all considered factors have significant effects at the 5\% level:
 
-Bar chart comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens SNR estimating HR - vs subject skin type on x axis.
+\begin{itemize}
+	\item \texttt{camera\_stationary} has a weak negative effect, meaning that estimations were slightly less accurate when camera was stationary,
+	\item \texttt{illuminance\_var} has a strong negative effect, meaning that estimations were less accurate when participant illuminance varied more, and
+	\item \texttt{movement} had a strong negative effect, meaning that higher participant movement led to less accurate estimations.
+\end{itemize}
 
-\subsection{Impact of subject age}
+The factors are collectively able to explain 6.8\% of the variance of the respiratory wave estimation SNR.
+As expected, participant movement degrades the estimation quality for the respiratory rate.
+It is surprising that variation in the illuminance of the participant's face impacts the estimation to a similar degree.
+Also, the finding that estimations were more accurate when the camera was not stationary is counterintuitive.
+One possible explanation is that VitalLens learns to understand that for some participants, repetitive movement of the (handheld) camera actually reveals information about respiration.
 
-Bar chart comparing G, CHROM, POS, DeepPhys, MTTS-CAN, VitalLens SNR estimating HR - vs subject age on x axis.
+Based on the results for both pulse and respiration, we advise users of VitalLens that for best performance they should (i) hold still, and (ii) make sure the face is evenly illuminated.
 
-\subsection{Impact of vitals}
+\subsubsection{Impact of participant movement}
 
-Investigate whether predictions are worse for lower / higher HR
+\pgfplotstableread[col sep=comma]{data/impact_movement.csv}\datatable
+
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+        ylabel=SNR,
+        height=5cm,
+        width=.8\textwidth,
+        legend style={
+            at={(0.98,.96)},
+            anchor=north east,
+            legend columns=-1
+        },
+        ybar,
+        bar width=14.,
+        xtick=data,
+        xticklabels from table={\datatable}{bin},
+        xticklabel style={rotate=30, anchor=east, align=center, text width=3cm},
+    ]
+    \addplot+[
+    	red!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=pulse_snr_mean, y error plus=pulse_snr_sd] {\datatable};
+    \addlegendentry{Pulse};
+    \addplot+[
+        blue!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=resp_snr_mean, y error plus=resp_snr_sd] {\datatable};
+    \addlegendentry{Resp};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Signal-to-noise ratio for pulse and respiration on the PROSIT test set, grouped by different levels of participant movement.}
+  \label{fig:impact-movement}
+\end{figure}
+
+According to the regression analysis, increasing amounts of participant movement lead to worse estimation performances for both heart rate and respiratory rate.
+We explore this effect in Figure \ref{fig:impact-movement} by looking the average SNR for different buckets of participant movement on the PROSIT test set.
+It is evident that SNR values indeed decrease with increased movement.
+There appears to be a large drop-off especially going from chunks that have no movement to those that have a few movements.
+
+\subsubsection{Impact of participant illuminance variation}
+
+\pgfplotstableread[col sep=comma]{data/impact_illuminance.csv}\datatable
+
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+        ylabel=SNR,
+        height=5cm,
+        width=.8\textwidth,
+        legend style={
+            at={(0.98,.96)},
+            anchor=north east,
+            legend columns=-1
+        },
+        ybar,
+        bar width=14.,
+        xtick=data,
+        xticklabels from table={\datatable}{bin},
+        xticklabel style={rotate=30, anchor=east, align=center, text width=3cm},
+    ]
+    \addplot+[
+    	red!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=pulse_snr_mean, y error plus=pulse_snr_sd] {\datatable};
+    \addlegendentry{Pulse};
+    \addplot+[
+        blue!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=resp_snr_mean, y error plus=resp_snr_sd] {\datatable};
+    \addlegendentry{Resp};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Signal-to-noise ratio for pulse and respiration on the PROSIT test set, grouped by different levels of participant illuminance variation.}
+  \label{fig:impact-illuminance}
+\end{figure}
+
+The regression analysis showed that increasing amounts of variation in participant illuminance also led to worse estimation performances for both heart rate and respiratory rate.
+We explore this effect in Figure \ref{fig:impact-illuminance} by looking at the average SNR for different buckets of variation in participant illuminance on the PROSIT test set.
+This figure reveals that especially when going from chunks with moderate illuminance variation to those with extreme illuminance variation, we see a large drop-off in estimation accuracy.
+
+\subsubsection{Impact of participant age}
+
+\pgfplotstableread[col sep=comma]{data/impact_age.csv}\datatable
+
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+        x tick label style={
+            /pgf/number format/1000 sep=},
+        ylabel=SNR,
+        height=5cm,
+        width=.8\textwidth,
+        legend style={
+            at={(0.02,.96)},
+            anchor=north west,
+            legend columns=-1
+        },
+        ymin=0,
+        ybar,
+        bar width=10.,
+        xtick=data,
+        xticklabels from table={\datatable}{bin},
+        xticklabel style={rotate=22.5, anchor=east},
+    ]
+    \addplot+[
+    	red!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=pulse_snr_mean, y error plus=pulse_snr_sd] {\datatable};
+    \addlegendentry{Pulse};
+    \addplot+[
+        blue!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=resp_snr_mean, y error plus=resp_snr_sd] {\datatable};
+    \addlegendentry{Resp};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Signal-to-noise ratio for pulse and respiration on VV-Medium, grouped by different participant ages.}
+  \label{fig:impact-age}
+\end{figure}
+
+Since the regression results indicated that VitalLens is slightly more accurate overall for older participants, we explore this effect in more detail as well.
+Figure \ref{fig:impact-age} indeed shows that especially for participants under 20 years of age, the SNR for both pulse and respiration is tends to be lower.
+
+\subsubsection{Impact of participant skin type}
+
+\pgfplotstableread[col sep=comma]{data/impact_skin_type.csv}\datatable
+
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+        ylabel=SNR,
+        height=5cm,
+        width=.8\textwidth,
+        legend style={
+            at={(0.98,.96)},
+            anchor=north east,
+            legend columns=-1
+        },
+        ybar,
+        bar width=14.,
+        xtick=data,
+        xticklabels from table={\datatable}{bin},
+        xticklabel style={align=center, text width=3cm},
+    ]
+    \addplot+[
+    	red!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=pulse_snr_mean, y error plus=pulse_snr_sd] {\datatable};
+    \addlegendentry{Pulse};
+    \addplot+[
+        blue!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=resp_snr_mean, y error plus=resp_snr_sd] {\datatable};
+    \addlegendentry{Resp};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Signal-to-noise ratio for pulse and respiration on VV-Medium, grouped by different participant skin types.}
+  \label{fig:impact-skin-type}
+\end{figure}
+
+According to the regression analysis on VV-Medium, the pulse wave estimation had a lower accuracy especially for participants with skin type 6.
+Here, we seek to further explore this effect.
+We start by grouping the chunks in VV-Medium by skin type and then plot the average estimation SNR for both pulse wave and respiration wave estimation in Figure \ref{fig:impact-skin-type}.
+Indeed, we observe that consistent with our previous regression analysis, there is a drop-off in the SNR between skin type 5 and 6 for pulse, while there is no obvious effect visible for respiration.
+
+\pgfplotstableread[col sep=comma]{data/impact_skin_type_comparison.csv}\datatable
+
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+        ylabel=SNR,
+        height=5cm,
+        width=.8\textwidth,
+        legend style={
+            at={(0.98,.96)},
+            anchor=north east,
+            legend columns=-1
+        },
+        ybar,
+        bar width=14.,
+        xtick=data,
+        xticklabels from table={\datatable}{bin},
+        xticklabel style={align=center, text width=3cm},
+    ]
+    \addplot+[
+    	red!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=pulse_snr_mean_vl, y error plus=pulse_snr_sd_vl] {\datatable};
+    \addlegendentry{VitalLens};
+    \addplot+[
+        green!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=pulse_snr_mean_vl_64, y error plus=pulse_snr_sd_vl_64] {\datatable};
+    \addlegendentry{VitalLens*};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Comparing signal-to-noise ratio between VitalLens and VitalLens* on VV-Medium, grouped by participant skin types.
+VitalLens* was solely trained on PROSIT, while VitalLens* was trained on both PROSIT and VV-Ghana-Small.}
+  \label{fig:impact-skin-type-comparison}
+\end{figure}
+
+Next, we want to investigate to what degree the inclusion of VV-Ghana-Small in our training data lessened this skin type effect.
+For this purpose, we compare VitalLens with \textit{VitalLens*}, a version of VitalLens that is solely trained on the PROSIT test set.
+The training set of PROSIT by itself only includes 2 individuals of skin type 6, a number that is massively increased in the combined training data.
+Both VitalLens and VitalLens* are then evaluated on VV-Medium.
+Figure \ref{fig:impact-skin-type-comparison} compares the results, again grouped by skin type.
+It is evident that the addition of VV-Ghana-Small led to a sizeable improvement mainly for skin types 5 and 6, but also skin types 3 and 4.
+
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+        ylabel=SNR,
+        height=5cm,
+        width=.8\textwidth,
+        legend style={
+            at={(0.98,.96)},
+            anchor=north east,
+            legend columns=-1
+        },
+        ybar,
+        bar width=14.,
+        xtick=data,
+        xticklabels from table={\datatable}{bin},
+        xticklabel style={align=center, text width=3cm},
+    ]
+    \addplot+[
+    	red!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=pulse_snr_mean_vl, y error plus=pulse_snr_sd_vl] {\datatable};
+    \addlegendentry{VitalLens};
+    \addplot+[
+        green!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=pulse_snr_mean_pos, y error plus=pulse_snr_sd_pos] {\datatable};
+    \addlegendentry{POS};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Comparing signal-to-noise ratio between VitalLens and POS on VV-Medium, grouped by participant skin types.}
+  \label{fig:impact-skin-type-comparison-pos}
+\end{figure}
+
+Finally, we compare the impact of skin types on pulse wave estimation between our learned model VitalLens and the best-performing handcrafted model POS.
+As visible in Figure \ref{fig:impact-skin-type-comparison-pos}, the impact of skin types in POS is even worse than on VitalLens*, with an even larger drop-off in SNR between skin types 5 and 6.
+
+We can conclude from this analysis that issues arising from the lower pulse signal strength in rPPG for darker skin types can at least partially be overcome by learning-based rPPG models which use training data that actually includes a significant amount of individuals with those skin types.
+
+\subsubsection{Impact of participant vitals}
+
+Finally, we investigate the impact of the vitals themselves on estimation performance.
+Although we took steps to prevent this, the distribution of vitals in the training set may have biased VitalLens towards predicting the average values found in the training set.
+The regressions found that estimations generally became slightly less accurate for individuals with higher heart rates and respiratory rates.
+Figures \ref{fig:impact-hr} and \ref{fig:impact-rr} show the SNR for pulse and respiration estimation, grouped by different heart rate and respiratory rates respectively.
+
+\pgfplotstableread[col sep=comma]{data/impact_hr.csv}\datatable
+
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+        x tick label style={
+            /pgf/number format/1000 sep=},
+        ylabel=SNR,
+        height=5cm,
+        width=.8\textwidth,
+        legend style={
+            at={(0.02,.96)},
+            anchor=north west,
+            legend columns=-1
+        },
+        ybar,
+        bar width=12.,
+        xtick=data,
+        xticklabels from table={\datatable}{bin},
+        xticklabel style={rotate=22.5, anchor=east},
+    ]
+    \addplot+[
+    	red!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=pulse_snr_mean, y error plus=pulse_snr_sd] {\datatable};
+    \addlegendentry{Pulse};
+    \addplot+[
+        blue!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=resp_snr_mean, y error plus=resp_snr_sd] {\datatable};
+    \addlegendentry{Resp};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Signal-to-noise ratio for pulse and respiration on VV-Medium, grouped by different participant heart rate values.}
+  \label{fig:impact-hr}
+\end{figure}
+
+\pgfplotstableread[col sep=comma]{data/impact_rr.csv}\datatable
+
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+        x tick label style={
+            /pgf/number format/1000 sep=},
+        ylabel=SNR,
+        height=5cm,
+        width=.8\textwidth,
+        legend style={
+            at={(0.02,.96)},
+            anchor=north west,
+            legend columns=-1
+        },
+        ybar,
+        bar width=14.,
+        xtick=data,
+        xticklabels from table={\datatable}{bin},
+        xticklabel style={rotate=22.5, anchor=east},
+    ]
+    \addplot+[
+    	red!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=pulse_snr_mean, y error plus=pulse_snr_sd] {\datatable};
+    \addlegendentry{Pulse};
+    \addplot+[
+        blue!60,
+        ybar,
+        error bars/.cd,
+        y dir=both,
+        y explicit,
+    ] table[x expr=\coordindex, y=resp_snr_mean, y error plus=resp_snr_sd] {\datatable};
+    \addlegendentry{Resp};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Signal-to-noise ratio for pulse and respiration on VV-Medium, grouped by different participant respiratory rate values.}
+  \label{fig:impact-rr}
+\end{figure}
+
+The figures show that generally, VitalLens does not appear to be biased by the distributions of heart rate and respiratory rate in the training data.
+We do however note that pulse estimation became less accurate for the chunks with the highest heart rates, and respiration estimation became less accurate for the chunks with the highest respiration rates.
+This consistent with the findings of the regression analysis.
 
@@
 
-Authors may wish to optionally include extra information (complete proofs, additional experiments and plots) in the appendix.
-All such materials should be part of the supplemental material (submitted separately) and should NOT be included in the main submission.
+VitalLens emerges as a novel solution in the domain of rPPG, showcasing its potential for accurate and real-time estimation of vital signs from selfie videos.
+The extensive evaluation establishes its superiority over established methods, such as G, CHROM, POS, DeepPhys, and MTTS-CAN.
+Notably, VitalLens outperforms these methods while maintaining fast inference times.
+The thorough analysis of factors impacting performance reveals the significance of participant movement and illuminance variation, providing valuable insights for optimal usage.
+Additionally, the inclusion of diverse training data to address skin type bias, contributes to the model's robustness.
 

### 2023-12-11 paper.py

@@ -140,6 +140,14 @@ raw_pos_prosit = pd.read_csv("raw_data/evaluate_pos_prosit.csv")
 raw_pos_prosit = raw_pos_prosit[['chunk_id', 'gt_hr_pox', 'gt_rr', 'pred_hr', 'pred_rr', 'conf_pulse_mean', 'conf_resp_mean', 'pulse_mae', 'pulse_mse', 'pulse_rmse', 'pulse_cor', 'pulse_snr', 'hr_ae', 'resp_mae', 'resp_mse', 'resp_rmse', 'resp_cor', 'resp_snr', 'rr_ae', 'live_error', 'live_error_t']]
 raw_pos_prosit = raw_pos_prosit.rename(columns={'chunk_id': 'id'})
 results_prosit_test = merge_force_suffix(results_prosit_test, raw_pos_prosit, on='id', how='inner', suffixes=['', '_pos'])
+raw_deep_phys_prosit = pd.read_csv("raw_data/evaluate_deep_phys_prosit.csv")
+raw_deep_phys_prosit = raw_deep_phys_prosit[['chunk_id', 'gt_hr_pox', 'gt_rr', 'pred_hr', 'pred_rr', 'conf_pulse_mean', 'conf_resp_mean', 'pulse_mae', 'pulse_mse', 'pulse_rmse', 'pulse_cor', 'pulse_snr', 'hr_ae', 'resp_mae', 'resp_mse', 'resp_rmse', 'resp_cor', 'resp_snr', 'rr_ae', 'live_error', 'live_error_t']]
+raw_deep_phys_prosit = raw_deep_phys_prosit.rename(columns={'chunk_id': 'id'})
+results_prosit_test = merge_force_suffix(results_prosit_test, raw_deep_phys_prosit, on='id', how='inner', suffixes=['', '_deepphys'])
+raw_mtts_can_prosit = pd.read_csv("raw_data/evaluate_mtts_can_prosit_8.csv")
[…]

### 2023-12-11 paper.tex
Changes based on Pieter's feedback; Version submitted to arXiv

@@
 
-\usepackage[final]{assets/techreport}
-%\usepackage[nonatbib]{assets/techreport}
+\usepackage[final,nonatbib]{assets/techreport}
 
@@
 
-\csvreader[before reading=\def\totalparticipantsvvghana{0}\def\totalchunksvvghana{0}\def\totaltimevvghana{0}]{data/vv_ghana_small_summary.csv}{participants=\participants,chunks=\chunks,time=\time}{%
-	\pgfmathsetmacro{\totalparticipantsvvghana}{\totalparticipantsvvghana+\participants}%
-	\pgfmathsetmacro{\totalchunksvvghana}{\totalchunksvvghana+\chunks}%
-	\pgfmathsetmacro{\totaltimevvghana}{\totaltimevvghana+\time}%
+\csvreader[before reading=\def\totalparticipantsvvafrica{0}\def\totalchunksvvafrica{0}\def\totaltimevvafrica{0}]{data/vv_africa_small_summary.csv}{participants=\participants,chunks=\chunks,time=\time}{%
+	\pgfmathsetmacro{\totalparticipantsvvafrica}{\totalparticipantsvvafrica+\participants}%
+	\pgfmathsetmacro{\totalchunksvvafrica}{\totalchunksvvafrica+\chunks}%
+	\pgfmathsetmacro{\totaltimevvafrica}{\totaltimevvafrica+\time}%
 }
@@
 Video of the human face and upper body has proven to be a rich source of information about an individual's physiological state \cite{verkruysse2008remote}.
-In particular, signals embedded in these videos can be harnessed to estimate vital signs such as heart rate and respiratory rate through a process known as Remote Photoplethysmography (rPPG).
+In particular, signals embedded in these videos can be harnessed to estimate vital signs through a process known as Remote Photoplethysmography (rPPG).
 This capability holds immense potential for non-invasive and real-time health monitoring applications.
@@
 Handcrafted algorithms offer advantages of fast inference speed and no training data required, but usually lack in accuracy; notable approaches include the original approach G \cite{verkruysse2008remote}, and the more sophisticated CHROM \cite{de2013robust} and POS \cite{wang2017algorithmic}.
-Given enough high-quality data to learn from, learning-based approaches such as DeepPhys \cite{chen2018deep} and MTTS-CAN \cite{liu2020multi} hold the promise of greater accuracy, which they have to trade off against inference speed.
-There are also other learning-based approaches achieving higher accuracy, which we do not cover here as they are not aiming to achieve real-time inference.
+Given enough high-quality data to learn from, learning-based approaches such as DeepPhys \cite{chen2018deep} and MTTS-CAN \cite{liu2020multi} hold the promise of greater accuracy, which is traded off against inference speed.
+There are also other learning-based approaches, which we do not cover here as they are not aiming to achieve real-time inference.
 
@@
 \begin{itemize}
-	\item \textbf{Real-time rPPG application.} We introduce VitalLens -- the first widely distributed rPPG application for real time estimation of vitals such as of heart rate and respiratory rate. 
+	\item \textbf{Real-time rPPG application.} We introduce VitalLens -- the first widely distributed rPPG application for real time estimation of vitals such as of heart rate and respiratory rate.\footnote{VitalLens is currently in private beta test and/or under review by Apple.
+This will be updated with a public App Store link once available.}
 	\item \textbf{Comprehensive evaluation on a large dataset.} We evaluate the performance of VitalLens on multiple datasets, including the large VV-Medium \cite{toye2023vital} with \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvmedium} participants.
@@
 
-VitalLens is trained on the \textit{PROSIT} and \textit{Vital Videos Ghana} datasets.
+VitalLens is trained on the \textit{PROSIT} and \textit{Vital Videos Africa} datasets.
 The vast majority of training data comes from PROSIT.
-For evaluation, we use the entire \textit{Vital Videos Medium} dataset as well as the test sets of the \textit{PROSIT} and \textit{Vital Videos Ghana} datasets.
+For evaluation, we use the entire \textit{Vital Videos Medium} dataset as well as the test sets of the \textit{PROSIT} and \textit{Vital Videos Africa} datasets.
 
@@
 
-\subsection{Vital Videos Ghana}
-
-Vital Videos Ghana is a new dataset from the authors of Vital Videos aiming to address the insufficient share of participants with skin types 5 and 6 usually found in rPPG datasets.
+\subsection{Vital Videos Africa}
+
+Vital Videos Africa is a new dataset from the authors of Vital Videos aiming to address the insufficient share of participants with skin types 5 and 6 usually found in rPPG datasets.
 It was collected in Ghana, with the majority of participants having skin type 5 or 6.
-We use a small instantiation (``VV-Ghana-Small''), which consists of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvghana} participants.
-Both camera and participants are stationary in this dataset.
+We use a small instantiation (``VV-Africa-Small''), which consists of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvafrica} participants.
+Both camera and participants are stationary in this subset.
 
@@
 
-We pre-process VV-Ghana-Small using the same steps applied to PROSIT.
+We pre-process VV-Africa-Small using the same steps applied to PROSIT.
 As part of this step, we create small chunks, calculate summary vitals, and extract further metadata.
@@
 \begin{table}[h!]
- 	\caption{VV-Ghana-Small Dataset Size}
- 	\label{tab:vv-ghana-small-summary}
+ 	\caption{VV-Africa-Small Dataset Size}
+ 	\label{tab:vv-africa-small-summary}
  	\centering
@@
     	table head=\toprule Split & \# Participants & \# Chunks & Time \\ \midrule,
-    	table foot=\midrule Total & \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvghana} & \sisetup{round-mode=places,round-precision=0}\num{\totalchunksvvghana} & \sisetup{round-mode=places,round-precision=1}\SI{\totaltimevvghana}{\hour} \\ \bottomrule,
+    	table foot=\midrule Total & \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvafrica} & \sisetup{round-mode=places,round-precision=0}\num{\totalchunksvvafrica} & \sisetup{round-mode=places,round-precision=1}\SI{\totaltimevvafrica}{\hour} \\ \bottomrule,
     	late after line=\\,
     	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
-  	]{data/vv_ghana_small_summary.csv}{}%
+  	]{data/vv_africa_small_summary.csv}{}%
   {\csvcoli & \num{\csvcolii} & \num{\csvcoliii} & \sisetup{round-mode=places,round-precision=1}\SI{\csvcoliv}{\hour} }
@@
 
-\subsection{Final training dataset: PROSIT + VV-Ghana-Small}
+\subsection{Final training dataset: PROSIT + VV-Africa-Small}
 \csvreader[before reading=\def\totalparticipantstraining{0}\def\totalchunkstraining{0}\def\totaltimetraining{0}]{data/training_summary.csv}{participants=\participants,chunks=\chunks,time=\time}{%
@@
 	}
-We combine the training sets of PROSIT and VV-Ghana-Small for training.
+We combine the training sets of PROSIT and VV-Africa-Small for training.
 As of this writing, this makes a total of \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantstraining} unique participants, \sisetup{round-mode=places,round-precision=0}\num{\totalchunkstraining} chunks or \sisetup{round-mode=places,round-precision=1}\num{\totaltimetraining} hours of data.
@@
 While developing the model, we train many different versions of our model.
-During training, we continuously monitor the model's performance on the validation sets of PROSIT and VV-Ghana-Small as outlined in Section \ref{sec:datasets}.
+During training, we continuously monitor the model's performance on the validation sets of PROSIT and VV-Africa-Small as outlined in Section \ref{sec:datasets}.
 Since the training dataset is disjoint from both validation sets in terms of participants, this allows us to calculate validation metrics which measure how well each model has learned to generalize in estimating vital signs from video.
@@
 	\item The test set of the PROSIT dataset. This is important as PROSIT includes camera and participant movement, which is to be expected when VitalLens is used in the wild.
-	\item The test set of the VV-Ghana-Small dataset.
+	\item The test set of the VV-Africa-Small dataset.
 \end{itemize}
@@
 
-We evaluated VitalLens on three datasets - VV-Medium, the PROSIT test set, and the VV-Ghana-Small test set.
+We evaluated VitalLens on three datasets - VV-Medium, the PROSIT test set, and the VV-Africa-Small test set.
 None of the participants included in these datasets were previously seen by VitalLens.
@@
 Results for the PROSIT test set are reported in Table \ref{tab:results-prosit-test} -- again, VitalLens beats all listed methods for all considered evaluation metrics.
-However, we observe that here, the absolute errors achieved by VitalLens for heart rate and respiratory rate are 4.62 bpm and 3.76, respectively -- markedly higher than for VV-Medium.
+However, we observe that here, the absolute errors achieved by VitalLens for heart rate and respiratory rate are 4.62 bpm and 3.76 bpm, respectively -- markedly higher than for VV-Medium.
 This is a first hint that factors like participant and camera movement are the main drivers of lower estimation accuracies.
@@
 \begin{table}[h!]
- 	\caption{Vitals estimation results on VV-Ghana-Small test set}
- 	\label{tab:results-vv-ghana-small-test}
+ 	\caption{Vitals estimation results on VV-Africa-Small test set}
+ 	\label{tab:results-vv-africa-small-test}
  	\centering
@@
     	before reading={\catcode`\%=12}, after reading={\catcode`\%=14}
-  	]{data/results_vv_ghana_test.csv}{}%
+  	]{data/results_vv_africa_test.csv}{}%
   {\csvcoli & \printMetric{\csvcolii} & \printMetric{\csvcoliii} & \printMetric{\csvcoliv} & \printMetric{\csvcolv} & \printMetric{\csvcolvi} & \printMetric{\csvcolvii} }
@@
 
-Finally, Table \ref{tab:results-vv-ghana-small-test} lists the results for the VV-Ghana-Small test set.
+Finally, Table \ref{tab:results-vv-africa-small-test} lists the results for the VV-Africa-Small test set.
 The results paint a similar picture as the two previous datasets, with VitalLens beating out the other methods.
-In comparison to VV-Medium, we observe that the errors are generally higher - in this case, this is caused by the demographic differences between the participants.
+In comparison to VV-Medium, we observe that the errors are generally higher - in this case, this may be caused by the demographic differences between the participants and more challenging lighting conditions.
 We further investigate this in the following sections.
@@
 
-Overall, the results confirm that the behavioral factors of participant movement and variation in the illuminance of the participant's face dominate the results with have a significant negative impact on estimation performance.
+Overall, the results confirm that the behavioral factors of participant movement and variation in the illuminance of the participant's face dominate the results with a significant negative impact on estimation performance.
 It is interesting to note that at least for these datasets, the effect of illuminance variation seems to have a greater impact than participant movement.
@@
   \caption{Comparing signal-to-noise ratio between VitalLens and VitalLens* on VV-Medium, grouped by participant skin types.
-VitalLens* was solely trained on PROSIT, while VitalLens* was trained on both PROSIT and VV-Ghana-Small.}
+VitalLens* was solely trained on PROSIT, while VitalLens was trained on both PROSIT and VV-Africa-Small.}
   \label{fig:impact-skin-type-comparison}
@@
 
-Next, we want to investigate to what degree the inclusion of VV-Ghana-Small in our training data lessened this skin type effect.
+Next, we want to investigate to what degree the inclusion of VV-Africa-Small in our training data lessened this skin type effect.
 For this purpose, we compare VitalLens with \textit{VitalLens*}, a version of VitalLens that is solely trained on the PROSIT test set.
@@
 Figure \ref{fig:impact-skin-type-comparison} compares the results, again grouped by skin type.
-It is evident that the addition of VV-Ghana-Small led to a sizeable improvement mainly for skin types 5 and 6, but also skin types 3 and 4.
+It is evident that the addition of VV-Africa-Small led to a sizeable improvement mainly for skin types 5 and 6, but also skin types 3 and 4.
 
@@
 
-\section*{References}
-
 \bibliographystyle{plain}

### 2023-12-11 paper.py

@@ -13,10 +13,10 @@ subjects_prosit = sessions_prosit.drop_duplicates(subset='subject_id')
 chunks_vv = pd.read_csv("raw_data/export_vv_medium.csv")[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'env_illuminance', 'env_illuminance_d', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate']]
 chunks_vv['subject_id'] = chunks_vv['session_description'].str.split('/').str.get(1)
 sessions_vv = chunks_vv.drop_duplicates(subset='session_id')
-# VV-Ghana
-chunks_vv_ghana = pd.read_csv("raw_data/export_vv_ghana_small.csv")[['id', 'camera_stationary', 'duration', 'env_complexity', 'env_variability', 'env_illuminance', 'env_illuminance_d', 'subject_age', 'subject_facemask', 'subject_gaze', 'subject_gender', 'subject_glasses', 'subject_headwear', 'subject_illuminance_d', 'subject_movement', 'subject_skin_type', 'frame_avg_bp_dia', 'frame_avg_bp_sys', 'frame_avg_hr_pox', 'frame_avg_hrv_sdnn_ecg', 'frame_avg_rr', 'frame_avg_spo2', 'session_description', 'session_id', 'session_split', 'session_video_framerate']]
-chunks_vv_ghana['subject_id'] = chunks_vv_ghana['session_description'].str.split('/').str.get(1)
-sessions_vv_ghana = chunks_vv_ghana.drop_duplicates(subset='session_id')
+# VV-Africa
[…]

### 2024-01-03 paper.tex
Paper revision

@@
 VitalLens outperforms several existing methods including POS and MTTS-CAN on all datasets while maintaining a fast inference speed.
-On VV-Medium, VitalLens achieves absolute errors of 0.71 bpm for heart rate estimation, and 0.76 rpm for respiratory rate estimation.
+On VV-Medium, VitalLens achieves mean absolute errors of 0.71 bpm for heart rate estimation, and 0.76 bpm for respiratory rate estimation.
 \end{abstract}
@@
 This will be updated with a public App Store link once available.}
-	\item \textbf{Comprehensive evaluation on a large dataset.} We evaluate the performance of VitalLens on multiple datasets, including the large VV-Medium \cite{toye2023vital} with \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvmedium} participants.
+	\item \textbf{Comprehensive evaluation on a large dataset.} We evaluate the performance of VitalLens on multiple datasets, including VV-Medium \cite{toye2023vital} with \sisetup{round-mode=places,round-precision=0}\num{\totalparticipantsvvmedium} participants.
 VitalLens outperforms G, CHROM, POS, DeepPhys, and MTTS-CAN in terms of estimation accuracy.
-On VV-Medium, VitalLens achieves absolute errors of 0.71 bpm for heart rate estimation, and 0.76 rpm for respiratory rate estimation, while taking 18 ms to run inference.
+On VV-Medium, VitalLens achieves mean absolute errors of 0.71 bpm for heart rate estimation, and 0.76 bpm for respiratory rate estimation, while taking 18 ms to run inference.
 	\item \textbf{Analysis of factors impacting performance.} We systematically investigate factors impacting estimation performance.
@@
 
-\section{Model Architecture}
-\label{sec:scope}
+\section{Architecture}
+\label{sec:architecture}
+
+VitalLens takes a series of video frames as inputs, and uses them to estimate the pulse waveform and the respiration waveform.
+The corresponding heart rate (HR) and respiratory rate (RR) estimates are derived from these waveforms using Fast Fourier Transform (FFT).
+Please refer to Figure \ref{fig:overview} for a broad overview of how VitalLens operates, along with the evaluation method used in Section \ref{sec:results}.
+
+\begin{figure}[h!]
+	\centering
+ 	\includegraphics[width=\textwidth]{figures/overview.pdf}
+ 	\caption{VitalLens estimates vital sign waveforms from video frames.
+The app displays the video feed with an overlay of the estimated pulse and respiration waveforms as well as the derived heart rate and respiration rate.
+We report evaluation metrics for all displayed quantities by comparing the estimations with gold-standard labels.}
+ 	\label{fig:overview}
+\end{figure}
 
@@
 The model architecture is broadly based on the \textit{EfficientNetV2} \cite{tan2021efficient} model family, enhanced with several improvements in architecture and model optimization to enable efficient training and inference in the rPPG domain.
-
 This report does not contain any further detail about the architecture of the model or training methodology - we reserve this for future publication.
@@
 Note that for most chunks, the missing respiratory signal was synthetically created.\footnote{This was done using an earlier version of our model trained on PROSIT.
-We then manually verified the correctness by visual inspection of both the label and video, and discarded bad labels.}
+We then manually verified the correctness by visual inspection of both the label and video.}
 
@@
             		ylabel=Frequency,
-            		xlabel={Respiratory rate [rpm]},
+            		xlabel={Respiratory rate [bpm]},
             		xmin=-2,
@@
 Vitals estimation is performed separately for each chunk at 30 frames per second for all compared methods.
-We use the dataset labels and model estimations to calculate several evaluation metrics for each chunk:
-Mean absolute error (MAE, smaller is better), signal-to-noise ratio (SNR, larger is better), and pearson correlation coefficient ($r$, larger is better).
-We then report the mean for each of these metrics across all chunks.
+As shown in Figure \ref{fig:overview}, we use the dataset labels and model estimations to calculate several evaluation metrics for each chunk:
+
+\begin{itemize}
+	\item \textbf{Absolute error} (AE, smaller is better): Used to measure the absolute difference between the estimated rate and the true rate.
+	\item \textbf{Signal-to-noise ratio} (SNR, larger is better): Measures the ratio of the true rate signal to the noise present in the estimated waveform in decibels.
+	\item \textbf{Pearson correlation coefficient} ($r$, larger is better): Measures the the linear correlation between the estimated waveform and the true waveform.
+\end{itemize}
+
+We then report the mean for each of these metrics across all chunks in the given dataset.
+As is common, we denote the mean AE as MAE as is common, but continue to use the abbreviations SNR and $r$ when reporting the mean SNR and mean $r$.
 
@@
 		tabular=lccccccc,
-    	table head=\toprule Method & \multicolumn{1}{c}{Heart rate} & \multicolumn{2}{c}{Pulse wave} & \multicolumn{1}{c}{Respiratory rate\tnote{a}} & \multicolumn{2}{c}{Respiration wave\tnote{a}} & Inference \\ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & speed in ms\tnote{b} \\ \midrule,
+    	table head=\toprule Method & \multicolumn{1}{c}{Heart rate} & \multicolumn{2}{c}{Pulse wave} & \multicolumn{1}{c}{Respiratory rate\tnote{a}} & \multicolumn{2}{c}{Respiration wave\tnote{a}} & Inference \\ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & MAE $\downarrow$ & SNR $\uparrow$ & $r \uparrow$ & time in ms\tnote{b} \\ \midrule,
     	table foot=\bottomrule,
@@
 Results for VV-Medium are reported in Table \ref{tab:results-vv-medium}.
-The heart rate and respiratory rate estimated by VitalLens across all chunks was off by 0.71 bpm and 0.76 rpm on average from the gold-standard labels, while taking 18 ms for inference.
-VitalLens beats all other listed methods for all considered evaluation metrics, while maintaining a competitive inference speed that is sufficient to run in real time at 30 fps.\footnote{The inference speeds are for a single video frame, excluding face detection but including all other computation, such as resizing.
-We measure the inference speeds using our own implementations as an average across 1,000 iterations on a workstation with Nvidia GeForce RTX 3080 and AMD Ryzen 7 3700X.}
+The heart rate and respiratory rate estimated by VitalLens across all chunks was off by 0.71 bpm and 0.76 bpm on average from the gold-standard labels, while taking 18 ms for inference.
+
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}{0.48\textwidth}
+    \begin{tikzpicture}
+      \begin{axis}[
+          xlabel=Gold-standard heart rate,
+          ylabel=Estimated heart rate,
+          height=7cm,
+          width=\textwidth,
+      ]
+      \addplot[
+          only marks,
+          mark=o,
+          color=red,
+      ] table [col sep=comma, x=hr_true, y=hr_est] {data/results_vv_medium_scatter.csv};
+      \addplot [gray,line width=0.2, domain=35:130, samples=2] {x};
+      \end{axis}
+    \end{tikzpicture}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.48\textwidth}
+    \begin{tikzpicture}
+      \begin{axis}[
+          xlabel=Gold-standard respiratory rate,
+          ylabel=Estimated respiratory rate,
+          height=7cm,
+          width=\textwidth,
+      ]
+      \addplot[
+          only marks,
+          mark=o,
+          color=blue,
+      ] table [col sep=comma, x=rr_est, y=rr_true] {data/results_vv_medium_scatter.csv};
+      \addplot [gray,line width=0.2, domain=4:44, samples=2] {x};
+      \end{axis}
+    \end{tikzpicture}
+  \end{subfigure}
+  \caption{VitalLens estimated vitals vs. gold-standard true vitals on VV-Medium. Besides a few outliers, estimations closely match the true vitals.}
+  \label{fig:two_scatterplots}
+\end{figure}
+
+VitalLens beats all other listed methods for all considered evaluation metrics, while maintaining a competitive inference speed that is sufficient to run in real time at 30 fps.\footnote{The inference times are for a single video frame, excluding face detection but including all other computation, such as resizing.
+We measure these times using our own implementations as an average across 1,000 iterations on a workstation with Nvidia GeForce RTX 3080 and AMD Ryzen 7 3700X.}
 The next best predictions for heart rate and respiratory rate also came from learning-based models MTTS-CAN and DeepPhys, respectively.
@@
         legend style={
-            at={(0.98,.96)},
-            anchor=north east,
+            at={(0.014,.04)},
+            anchor=south west,
             legend columns=-1
@@
     ]
+    \draw[gray, line width=0.2] ({rel axis cs:0,0}|-{axis cs:0,0}) -- ({rel axis cs:1,0}|-{axis cs:0,0});
     \addplot+[
@@
         y explicit,
-    ] table[x expr=\coordindex, y=pulse_snr_mean, y error plus=pulse_snr_sd] {\datatable};
+    ] table[x expr=\coordindex, y=pulse_snr_mean, y error=pulse_snr_sd] {\datatable};
     \addlegendentry{Pulse};
@@
         y explicit,
-    ] table[x expr=\coordindex, y=resp_snr_mean, y error plus=resp_snr_sd] {\datatable};
+    ] table[x expr=\coordindex, y=resp_snr_mean, y error=resp_snr_sd] {\datatable};
     \addlegendentry{Resp};
@@
         legend style={
-            at={(0.02,.96)},
+            at={(0.014,.96)},
             anchor=north west,
@@
         y explicit,
-    ] table[x expr=\coordindex, y=pulse_snr_mean_vl, y error plus=pulse_snr_sd_vl] {\datatable};
+    ] table[x expr=\coordindex, y=pulse_snr_mean_vl, y error=pulse_snr_sd_vl] {\datatable};
     \addlegendentry{VitalLens};
@@
         y explicit,
-    ] table[x expr=\coordindex, y=pulse_snr_mean_vl_64, y error plus=pulse_snr_sd_vl_64] {\datatable};
+    ] table[x expr=\coordindex, y=pulse_snr_mean_vl_64, y error=pulse_snr_sd_vl_64] {\datatable};
     \addlegendentry{VitalLens*};
@@
 Next, we want to investigate to what degree the inclusion of VV-Africa-Small in our training data lessened this skin type effect.
-For this purpose, we compare VitalLens with \textit{VitalLens*}, a version of VitalLens that is solely trained on the PROSIT test set.
+For this purpose, we compare VitalLens with \textit{VitalLens*}, a version of VitalLens that is solely trained on PROSIT.
 The training set of PROSIT by itself only includes 2 individuals of skin type 6, a number that is massively increased in the combined training data.
@@
         y explicit,
-    ] table[x expr=\coordindex, y=pulse_snr_mean_pos, y error plus=pulse_snr_sd_pos] {\datatable};
+    ] table[x expr=\coordindex, y=pulse_snr_mean_pos, y error=pulse_snr_sd_pos] {\datatable};
     \addlegendentry{POS};
@@
 As visible in Figure \ref{fig:impact-skin-type-comparison-pos}, the impact of skin types in POS is even worse than on VitalLens*, with an even larger drop-off in SNR between skin types 5 and 6.
+% TODO: Is this correct?
 
@@
         legend style={
-            at={(0.02,.96)},
-            anchor=north west,
+            at={(0.014,.04)},
+            anchor=south west,
             legend columns=-1
@@
         bar width=14.,
+        ymin=-7,
         xtick=data,
@@
 
+\begin{ack}
+We are thankful to Pieter-Jan Toye and Andrew Goodwin for the helpful suggestions.
+This work was funded by Rouast Labs Pty Ltd.
+\end{ack}
+
 \bibliographystyle{plain}

### 2024-01-03 paper.py

@@ -167,15 +167,15 @@ raw_pos_vv = raw_pos_vv[['chunk_id', 'pred_hr', 'pred_rr', 'conf_pulse_mean', 'c
 raw_pos_vv = raw_pos_vv.rename(columns={'chunk_id': 'id'})
 results_vv = merge_force_suffix(results_vv, raw_pos_vv, on='id', how='inner', suffixes=['', '_pos'])
 raw_deep_phys_vv = pd.read_csv("raw_data/evaluate_deep_phys_vv.csv")
-raw_deep_phys_vv = raw_deep_phys_vv[['chunk_id', 'gt_hr_pox', 'gt_rr', 'pred_hr', 'pred_rr', 'conf_pulse_mean', 'conf_resp_mean', 'pulse_mae', 'pulse_mse', 'pulse_rmse', 'pulse_cor', 'pulse_snr', 'hr_ae', 'resp_mae', 'resp_mse', 'resp_rmse', 'resp_cor', 'resp_snr', 'rr_ae', 'live_error', 'live_error_t']]
+raw_deep_phys_vv = raw_deep_phys_vv[['chunk_id', 'pred_hr', 'pred_rr', 'conf_pulse_mean', 'conf_resp_mean', 'pulse_mae', 'pulse_mse', 'pulse_rmse', 'pulse_cor', 'pulse_snr', 'hr_ae', 'resp_mae', 'resp_mse', 'resp_rmse', 'resp_cor', 'resp_snr', 'rr_ae', 'live_error', 'live_error_t']]
 raw_deep_phys_vv = raw_deep_phys_vv.rename(columns={'chunk_id': 'id'})
 results_vv = merge_force_suffix(results_vv, raw_deep_phys_vv, on='id', how='inner', suffixes=['', '_deepphys'])
 raw_mtts_can_vv = pd.read_csv("raw_data/evaluate_mtts_can_vv_8.csv")
-raw_mtts_can_vv = raw_mtts_can_vv[['chunk_id', 'gt_hr_pox', 'gt_rr', 'pred_hr', 'pred_rr', 'conf_pulse_mean', 'conf_resp_mean', 'pulse_mae', 'pulse_mse', 'pulse_rmse', 'pulse_cor', 'pulse_snr', 'hr_ae', 'resp_mae', 'resp_mse', 'resp_rmse', 'resp_cor', 'resp_snr', 'rr_ae', 'live_error', 'live_error_t']]
[…]
